--- 
title       : "Experiments 11: Putting it All Together"
description : "This video and exercise wrap up what we have learned about experiments."
  
--- type:VideoExercise lang:r aspect_ratio:62.5 xp:50 skills:1 key:a39c1976d4
## Splitting Samples, Dimensionality, and Big Data
*** =video_link
//player.vimeo.com/video/230616790
 

--- type:NormalExercise lang:r aspect_ratio:62.5 xp:100 skills:2 key:5b941b8026
## Putting it All Together
Often, our experiments don't go as planned. This long and complex exercise will test your knowledge in resolving a question with realistic problems. 

KittyCatch is a location-based augmented reality game developed by Meowtec. In this game, players use their cell phones' GPS abilities to locate and capture virtual kittens. Although KittyCatch is free to play, Meowtec generates revenue from KittyCatch through pop-up ads, which appear about once for every half-mile the player walks while using the application, up to 5 times a day. Since Meowtec's revenue is highly dependent on how far people walk while playing KittyCatch, their primary goal for future updates to KittyCatch is to incentivize play.

One strategy to increase users' play time is to increase the distance between KittyCatch's "points of kinterest"; that is, increase the distance that users must travel to capture new kittens. Many players stop their sessions of KittyCatch before they catch a single kitten, and may be turned off by the need to travel farther to catch a kitten. However, Meowtec believes that the loss of play time for these players would be offset by the many other players who consistently travel as far as needed to catch a kitten. 

To test this hypothesis, Meowtec constructs an experiment: For a small random sample of players in Springfield, Massachusetts, Meowtec increases the distance that players must walk to reach a point of kinterest (specifically, from .25 miles to .5 miles). They compare these players to an equal random sample of players in Springfield, Massachusetts that have the default distance between points of kinterest (.25 miles). However, as they explore their data, they find a few problems with their experiment. 

With the dataset, `KittyCatch`, and the annotations in the R workspace, help Meowtec determine a valid and reliable average treatment effect from their experiment. Specifically, balance the data and correct for any coding errors in the original dataset to estimate a reliable average treatment effect. This question is much longer than previous questions, but also more representative of the types of steps that data scientists use when working with experimental data. 

*** =instructions
- 1-10) Balance the data.
- 11-13) Correct for coding errors in the data.
- 14-15) Determine the average treatment effect.

*** =pre_exercise_code
```{r}

library (dplyr)
library (data.table)
library (purr)

set.seed(1)
#Create rnorm function that allows for random distribution with min and max
  rtnorm <- function(n, mean, sd, min = -Inf, max = Inf){
    qnorm(runif(n, pnorm(min, mean, sd), pnorm(max, mean, sd)), mean, sd)
  }
#Create rounding function that allows to round to numbers above 1
  mround <- function(x,base){ 
    base*round(x/base) 
  } 

n=688
#Create dataframe
  KittyCatch<-as.data.frame(matrix(nrow=n,ncol=4))
  names(KittyCatch)<-c("DistanceWalked","Age","TotalHoursPlayed","Treatment")
#Assign Treatment randomly
  KittyCatch$Treatment<-rbinom(n,1,.35)
#Assign Values to variables
 #DistanceWalked
  #Create Variable
    KittyCatch$DistanceWalked<-round(rtnorm(n=n, mean=.62, sd=.3, min = 0, max = 1.3),2)
  #Assign 4 Outliers to DistanceWalked among control
    KittyCatch$DistanceWalked[KittyCatch$Treatment==0][sample(sum(KittyCatch$Treatment==0),4)]<-round(rtnorm(n=4, mean=7.00, sd=.6, min = 5, max = 9),2)
  #Age
    #Create Variable
      KittyCatch$Age<-round(rtnorm(n=n, mean=14, sd=4, min = 10, max = 40))
    #Assign age to be correlated with DistanceWalked
      KittyCatch$Age<-round(KittyCatch$Age+(ifelse(KittyCatch$DistanceWalked>median(KittyCatch$DistanceWalked),2,-2)))
    #Assign age to be correlated with treatment above age 18
      KittyCatch$Age<-round(KittyCatch$Age+(ifelse(KittyCatch$Age>17 & KittyCatch$Treatment==1,4,0)))
    #Top code in control group
      KittyCatch$Age[KittyCatch$Treatment==0]<-ifelse(KittyCatch$Age[KittyCatch$Treatment==0]>18,18,KittyCatch$Age[KittyCatch$Treatment==0])
  #TotalHoursPlayed
    #Create Variable
      KittyCatch$TotalHoursPlayed<-round(rtnorm(n=n, mean=30, sd=10, min = 1, max = 50))
    #Make it correlated with distance walked
      KittyCatch$TotalHoursPlayed<-round(.8*KittyCatch$TotalHoursPlayed + .2*KittyCatch$TotalHoursPlayed*KittyCatch$DistanceWalked)
    #Randomly assign missing values, make the proportion larger in the treatment; and make the likelihood for NA correlate with larger values
      KittyCatch$TotalHoursPlayed[KittyCatch$Treatment==0]<-ifelse(rbinom(sum(KittyCatch$Treatment==0),1,(.05+.05*KittyCatch$TotalHoursPlayed[KittyCatch$Treatment==0]/mean(KittyCatch$TotalHoursPlayed))),NA,KittyCatch$TotalHoursPlayed[KittyCatch$Treatment==0])
      KittyCatch$TotalHoursPlayed[KittyCatch$Treatment==1]<-ifelse(rbinom(sum(KittyCatch$Treatment==1),1,(.25+.05*KittyCatch$TotalHoursPlayed[KittyCatch$Treatment==1]/mean(KittyCatch$TotalHoursPlayed,na.rm = T))),NA,KittyCatch$TotalHoursPlayed[KittyCatch$Treatment==1])
  #Make distance walked negatively correlated with treatment prior to balance, and positvely correlated with treatment after balance
      removed<-ifelse(KittyCatch$Age>17 | is.na(KittyCatch$TotalHoursPlayed),1,0)
      KittyCatch$DistanceWalked<-round(ifelse(removed==1 & KittyCatch$Treatment==0,1.5*KittyCatch$DistanceWalked,ifelse(removed==0 & KittyCatch$Treatment==1,1.3*KittyCatch$DistanceWalked,KittyCatch$DistanceWalked)),2)

```
*** =sample_code
```{r}

# PART I: EXPLORING OUR DATA

# 1) Let's dig into the data. We first run the structure function, str(), to get a sense of the dataset. DistanceWalked is our outcome of interest, and Treatment is the key independent variable that we care about. TotalHoursPlayed refers to how many hours the user has put into KittyCatch in total.

#---- Question 1:------#
     Solution1<- 
#-----------------------#

# 2) Although it is bad practice, let's begin this exercise by first taking a naive estimate of the average treatment effect on DistanceWalked before exploring the dataset. Subtract the mean of the control group from the mean of the treatment group.

#---- Question 2: ------#
     Solution2<- 
#-----------------------#

# 3) Now let's do that again, but with a t-test.  
    
#---- Question 3: ------#
     Solution3<- 
#-----------------------#

# You will see that, based on the t-test results, it seems that there is substantial negative treatment effect; respondents walked less in the treatment than in the control. But is this finding reliable?

# 4) Now let's examine our outcome of interest, DistanceWalked, with the summary function. 

#---- Question 4: ------#
     Solution4<- 
#-----------------------#

# Notice that the maximum value is much larger than the mean, median, and 3rd quartile of the dataset. This outlier might bias our t-test results by violating its assumption that the data is normally distributed. 

# 5) Run a density plot on DistanceWalked to observe the long right tail in the distribution of our outcome of interest.

#---- Question 5: ------#
     Solution5<- 
#-----------------------#
  
# 6) Since the dataset is relatively small, let's use the `head` and `order` commands to examine the highest DistanceWalked observations in the dataset. The following syntax will order the top results by the outcome of interest, so use it as a guide: head(Dataframe[order(Dataframe$Variable,decreasing=T),])

#---- Question 6: ------#
     Solution6<- 
#-----------------------#

# 7) Notice that the four largest values for DistanceWalked are more than three times as large as the 5th largest value for DistanceWalked. These are clearly outliers. However, we should also note that all of these values are in the control group, so removing them from our dataset, as we had done with previous outliers, might also bias our estimates by lowering the mean values for the treatment group. Instead, let's say that our maximum distance we will allow is 2. This is known as top coding, and we'll top code these values so that they are still the largest values in the dataset, but so that they are not outliers and do not violate our assumptions of a normal distribution. Top coding can be arbitrary (and there are better methods for handling outliers), but for this assignment, use the following syntax as a model to create an ifelse statement that says that any value of DistanceWalked over 2 will be changed to equal 2: dataframe$variable<-ifelse(dataframe$variable>x,y,dataframe$variable). Note: x is the current value of the variable, and y is the final value we want.

#---- Question 7: ------#
     KittyCatch$DistanceWalked<-
#-----------------------#
    
# 8) If you completed question 7 correctly, the distribution of DistanceWalked should now look much more normal. Create a new density plot of DistanceWalked to see if that is true.

#---- Question 8: ------#
     Solution8<- 
#-----------------------#

# Nice work so far.


# PART II: CHECKING FOR BALANCE

# 9) Next, we'll want to make sure that our control groups and treatment groups are balanced. Let's compare the age of the control group and treatment group with boxplots. The model syntax for creating this boxplot is: boxplot(dataframe$variable[dataframe$Treatment==0], dataframe$variable[dataframe$Treatment==1], names=c("Control","Treatment")) 

#---- Question 9: ------#
     Solution9<- 
#-----------------------#

# Notice that the medians and interquartile ranges are very similar between the treatment and control groups, but that the largest values in the treatment group rest much higher than in the control group. It turns out that this did not happen by chance; by default, KittyCatch only requires respondents to enter their ages if they state that they are younger than 18; however, for the experiment, all respondents in the treatment group were required to enter their ages. 

#10) It could be problematic if the ages among the treatment group are different than those in the control group, especially if Age is correlated with DistanceWalked. Run a cor() function between Age and DistanceWalked to see if they are indeed correlated. The syntax for this will be cor(dataframe$variable1, dataframe$variable2):

#---- Question 10: ------#
     Solution10<- 
#-----------------------#

# 11) Are the treatment and control groups balanced on Age, or are they imbalanced? Let's run a t-test to see.  The syntax for this will be: t.test(dataframe$variable[dataframe$Treatment==0],dataframe$variable[dataframe$Treatment==1])

#---- Question 11: ------#
     Solution11<- 
#-----------------------#

# As you can see, they are imbalanced. 

# 12) There are two options to deal with this problem. One option is to ignore it, as we know that age was top coded in the control group, and may be the entire cause of the imbalance between the treatment and control group. However, what if people who opted into the treatment group tended to be older on average? This could bias our results. The best way to deal with this problem is with a different modeling technique (e.g. regression models). However, since we want to continue using t-tests for this module, a good way to deal with this problem is to subset the data to only use observations we can be sure are similarly aged. Since we can only be sure of the age of respondents in the treatment group for those younger than 18, let's subset the dataset to only include respondents who are 17 or younger. The syntax for this will be dataframe[dataframe$variable<=17,].
  
#---- Question 12: Subset KittyCatch to respondents who are 17 or younger ----#
    KittyCatch<-
#----------------------------------------------------------------------------#

# 13) If you completed question 12 correctly, Age among the treatment and control groups should now be balanced. Run the t-test from Question 11 again to see if the two groups are now balanced on Age.

#---- Question 13: ------#
     Solution13<- 
#-----------------------#

# Great! We've made a lot of progress, so let's look further at the quality of our data.

 
 # MISSING DATA PROBLEMS?        
    
# 14) Now let's look for balance in the last variable we know about the treatment and control groups, TotalHoursPlayed. Run the summary function on the variable TotalHoursPlayed.  Moreover, as the correlation function below shows, TotalHours played appears to be correlated with DistanceWalked. 

#---- Question 14: ------#
     Solution14<- 
#-----------------------#

# 15) Notice from the summary function that 127 values are labeled "NA". These cases are missing information, in large part because KittyCatch allows users to opt out of recording how many hours they play. This could be very problematic, as it indicates that we only have distance walked information for respondents who selected to do so. Let's find the correlation between TotalHoursPlayed and DistanceWalked the entire sample population. The syntax for this will be: cor(dataframe$variable1, dataframe$variable2, use="complete").

#---- Question 15: ------#
     Solution15<- 
#-----------------------#

# That doesn't look too good.
    
# 16) Even more worrisome, the NA values for TotalHoursPlayed appear at higher rates for respondents in the treatment group than in the control group. Check the average values ofTotalHoursPlayed in both treatment and control groups, in separate lines, using the mean(is.na(dataframe$variable)) function. Remember that you can select units within a dataframe by their Treatment value with the syntax variable[dataframe$Treatment==x]. 

#---- Question 16: ------#
     Solution16<- 
#-----------------------#

# 17) This isn't the end for our analysis, though. Let's look for some good news about our data within TotalHoursPlayed. One simple thing we can do is see if that variable is balanced across the treatment and control groups for the units that we actually do have information for, so run a t-test between the treatment and control groups on the variable TotalHoursPlayed.

#---- Question 17: ------#
     Solution17<- 
#-----------------------#

# While total hours played appears balanced among the treatment and control groups for those whose information is available, the fact that its missingness is correlated with DistanceWalked and is also patterned among the treatment and control groups is cause for concern. Specifically, this could suggest that respondents in the treatment group play more KittyCatch than those in the control group, and are likely to walk further with KittyCatch due to this. 
  
# 18) A simple, albeit imperfect, solution to this problem is listwise deletion; that is, removing respondents from our sample who did not report their total hours played. Subset KittyCatch to respondents who reported their total hours played. The syntax for this will be: dataframe<-KittyCatch[!is.na(dataframe$variable),].

#---- Question 18: Subset KittyCatch to respondents who reported their total hours played ----#
  KittyCatch<-
#--------------------------------------------------------------------------------------------#

# Whew! We've explored the data, dealt with the issues, and now we can finally calculate the treatment effect of our experiment!


# CALCULATE AN AVERAGE TREATMENT EFFECT

# 19) Now that we have balanced our data on all known dimensions, let's return the average treatment effect of walking farther on time spent playing KittyCatch by subtracting the means of our control group from our treatment group:

#---- Question 19: ----#
  Solution19<-
#--------------------------------------------------------------------------------------------------#

# Great job. But is that value statistically significant? We don't want to get too excited about it until we know that as well.

# 20) Finally, let's test if the ATE is significant by finding out if the p value of a t-test is less than .05, which we will use as our standard of significance:

#---- Question 20 -------#
  Solution20<-
#--------------------------------------------------------------------------------------------------#
  
```
*** =solution
```{r}
  KittyCatch$DistanceWalked<-ifelse(KittyCatch$DistanceWalked>2,2,KittyCatch$DistanceWalked)
  KittyCatch<-KittyCatch[KittyCatch$Age<=17,]
  KittyCatch<-KittyCatch[!is.na(KittyCatch$TotalHoursPlayed),]
  Solution1<- str(KittyCatch)
  Solution2<- mean(KittyCatch$DistanceWalked[KittyCatch$Treatment==1]) -  mean(KittyCatch$DistanceWalked[KittyCatch$Treatment==0])
  Solution3<- t.test(KittyCatch$DistanceWalked[KittyCatch$Treatment==1],KittyCatch$DistanceWalked[KittyCatch$Treatment==0])
  Solution4<- summary(KittyCatch$DistanceWalked)
  Solution5<- plot(density(KittyCatch$DistanceWalked))
  Solution6<- head(KittyCatch[order(KittyCatch$DistanceWalked,decreasing=T),])
  Solution8<- plot(density(KittyCatch$DistanceWalked))
  Solution9<- boxplot(KittyCatch$Age[KittyCatch$Treatment==0],KittyCatch$Age[KittyCatch$Treatment==1],names=c("Control","Treatment"))
  Solution10<-cor(KittyCatch$Age,KittyCatch$DistanceWalked)
  Solution11<- t.test(KittyCatch$Age[KittyCatch$Treatment==0],KittyCatch$Age[KittyCatch$Treatment==1])
  Solution13<- t.test(KittyCatch$Age[KittyCatch$Treatment==0],KittyCatch$Age[KittyCatch$Treatment==1])
  Solution14<- summary(KittyCatch$TotalHoursPlayed)
  Solution15<- cor(KittyCatch$TotalHoursPlayed,KittyCatch$DistanceWalked,use="complete")
  Solution16<-     mean(is.na(KittyCatch$TotalHoursPlayed[KittyCatch$Treatment==1]))
    mean(is.na(KittyCatch$TotalHoursPlayed[KittyCatch$Treatment==0]))
  Solution17<-  t.test(KittyCatch$TotalHoursPlayed[KittyCatch$Treatment==1],KittyCatch$TotalHoursPlayed[KittyCatch$Treatment==0])
  Solution18<- KittyCatch<-KittyCatch[!is.na(KittyCatch$TotalHoursPlayed),]
  Solution19<- mean(KittyCatch$DistanceWalked[KittyCatch$Treatment==1])-mean(KittyCatch$DistanceWalked[KittyCatch$Treatment==0])
  Solution20<-t.test(KittyCatch$DistanceWalked[KittyCatch$Treatment==1],KittyCatch$DistanceWalked[KittyCatch$Treatment==0])$p.value
```
*** =sct
```{r}
test_object("Solution4")
test_object("Solution5")
success_msg("Good work! Our naive estimate of the average treatment effect was opposite our estimate after we tried to balance the data. However, we had to ignore a large fraction of our observations to get to that conclusion. This is the difficulty of computing conditional average treatment effects with t-tests. However, this problem can be sidestepped with regression models. Our next module teaches how to use regression models in detail.")
```
