--- 
title       : "Experiments 10: Bounds Analysis"
description : "These videos and a related exercise discuss how to form bounds on causal effects in the presence of experimental noncompliance."


--- type:VideoExercise lang:r aspect_ratio:62.5 xp:50 skills:1 key:25526ab294
## Bounds Analysis for Missing Data
*** =video_link
//player.vimeo.com/video/199858153

---
## Let's Code: Computing Bounds Under Noncompliance
```yaml
type: VideoExercise
lang: r
xp: 50
skills: 1
key: 06215f9d20
video_link: "//player.vimeo.com/video/279737481"
```


--- type:NormalExercise lang:r xp:100 skills:1 key:a42175703d
## Offering a Higher Credit Card Limit: Computing Bounds Under Noncompliance
Your first calculations on the CreditCo data in Chapter 7 showed that noncompliance was an issue in the experiment. In the previous chapter, you examined how to resolve this issue using a natural experiment. This time, we have provided you with more data to examine, particularly the credit ratings for the individuals in both the treatment and control groups. But goal remains the same as before: does the offer of a higher credit limit decrease the number of late payments (defaults)? 

In this exercise, you will look at 2 alternative methods for dealing with the non-compliance issue.  One method will be a bounds analysis, and the other will be either 1) assuming random compliance, or 2) an Intention-to-Treat analysis. First, you compute the bounds of average treatment effect of opting into a credit line increase offer. Second, you calculate what the average treatment effect would be in one of the two other methods (or both, if you want to be complete!).

*** =instructions
- 1) Determine the ratio of people who accepted the offer.
- 2) Compute upper bound on the average treatment effect under non-compliant behavior.
- 3) Compute lower bound on the average treatment effect under non-compliant behavior.
- 4) Compare this with the results of an intention-to-treat analysis, under random compliance, or both!

*** =hint
- When the values of a variable are binary (0 or 1), you can calculate the proportion of 1s by simply taking the mean of the variable.
- Remember to be aware of selecting the correct subsample.

*** =pre_exercise_code
```{r}

library(data.table)
library(dplyr)
  set.seed(1)
  n             <- 9e3
  frac_treated  <- .5
  frac_female   <- .5656
  frac_white    <- .688

# Initialize dataframe
  CreditCo <- as.data.frame(matrix(0, ncol=11,nrow=n))
  colnames(CreditCo) <- c("id","offered","opt_in","FICO","age","female","race_white","default_pre","default_post","balance_pre","balance_post")

# Simulate baseline data
  CreditCo$id         <- seq(1,n,1)
  CreditCo$offered    <- as.integer(runif(n)<frac_treated)
  CreditCo$opt_in     <- rep.int(0,n)
  CreditCo$FICO       <- rnorm(n, mean=736, sd=300)
  CreditCo$age        <- sample(18:55, n, replace=T)
  CreditCo$female     <- as.integer(runif(n)<frac_female)
  CreditCo$race_white <- as.integer(runif(n)<frac_white)

# make FICO score intelligible
  CreditCo$FICO[CreditCo$FICO>850] <- 850
  CreditCo$FICO[CreditCo$FICO<300] <- 300
  CreditCo$FICO                    <- round(CreditCo$FICO)

# simulate pre-experiment default rate
  draw <- runif(n)
  xb   <- -1.82-0.2*CreditCo$FICO/100+.046*(CreditCo$FICO^2)/10000-5.1*CreditCo$female-7*CreditCo$race_white
  p    <- exp(xb)/(1+exp(xb))
  CreditCo$default_pre <- as.integer(draw<p)

# simulate pre-experiment balance level
  draw <- rnorm(n, mean=0, sd=0.5)
  CreditCo$balance_pre <- 8.5-0.2*CreditCo$FICO/100+.046*(CreditCo$FICO^2)/10000-0.15*CreditCo$female-0.85*CreditCo$race_white + draw
  CreditCo$balance_pre <- exp(CreditCo$balance_pre)

# Simulate noncompliance and opt-in behavior
  draw <- runif(sum(CreditCo$offered))
  xb   <- 6.75-1.2*CreditCo$FICO/100+.07*(CreditCo$FICO^2)/10000-2.1*CreditCo$female-2*CreditCo$race_white
  p    <- exp(xb)/(1+exp(xb))
  CreditCo$opt_in[CreditCo$offered==1] <- as.integer(draw<p[CreditCo$offered==1])

# Simulate post outcomes
  draw <- runif(n)
  xb   <- -1.82-0.2*CreditCo$FICO/100+.046*(CreditCo$FICO^2)/10000-5.1*CreditCo$female-7*CreditCo$race_white+4*CreditCo$offered*CreditCo$opt_in
  p    <- exp(xb)/(1+exp(xb))
  CreditCo$default_post <- as.integer(draw<p)

# balance
  draw <- rnorm(n, mean=0, sd=0.5)
  CreditCo$balance_post <- 8.5-0.2*CreditCo$FICO/100+.046*(CreditCo$FICO^2)/10000-0.15*CreditCo$female-0.85*CreditCo$race_white + 1*CreditCo$offered*CreditCo$opt_in + draw
  CreditCo$balance_post <- exp(CreditCo$balance_post)

# Remove elements and variables from environment
  rm(draw,frac_female,frac_treated,frac_white,n,p,xb)
  CreditCo <- CreditCo[,c("id","offered","opt_in","FICO","female","race_white","default_pre","default_post","balance_pre","balance_post")]
```

*** =sample_code
```{r}
# The dataset `CreditCo` is available in your workspace.

# 1) Determine the fraction of opt-in among those offered the increase.


# 2) Calculate the upper bound of the ATE by assuming that all non-opt-in customers have a late payment, and subtract the non-offered average from this value.

    Solution1<-

# 3) Calculate the lower bound of the ATE by assuming that all non-opt-in customers don't have a late payment, and subtract the non-offered average from this value.

    Solution2<-

# 4) Now compare this result with the findings from another method: an intention-to-treat analysis, under random compliance, or both! To calculate the intention-to-treat analysis ATE, compare units who received the offer and those who did not receive the offer. To calculate the ATE under random compliance, simply calculate a "naive ATE" by comparing treatment and control groups . 


# 5) Calculate the p values of ATEs from Question 4. How do they compare to the bounds analysis?


```

*** =solution
```{r}
mean(CreditCo$opt_in[CreditCo$offered==1])
Solution1 <- Solution1*mean(CreditCo$default_post[CreditCo$opt_in==1]) + (1-Solution1)*1 - mean(CreditCo$default_post[CreditCo$offered==0])
Solution2 <- Solution1*mean(CreditCo$default_post[CreditCo$opt_in==1]) + (1-Solution1)*0 - mean(CreditCo$default_post[CreditCo$offered==0])

```

*** =sct
```{r}
test_object("Solution1")
test_object("Solution2")
test_object("Solution3")

success_msg("Nice job! As you can see, we get slightly different answers based on the method we use. What did we learn from this experiment overall? Well, that depended on the technique you used to analyze it. When working with data, usually the best option is to learn the answers from multiple techniques so that you can have a more complete and nuanced interpretation of what really happened.")
```

--- type:VideoExercise lang:r aspect_ratio:62.5 xp:50 skills:1 key:0588056e73
## Experimental Design & RCTs
*** =video_link
//player.vimeo.com/video/209271489
