--- 
title       : "Experiments 6: Common Issues in Experiments"
description : "These videos and exercises will discuss common issues in carrying out experiments"

--- type:VideoExercise lang:r aspect_ratio:62.5 xp:50 skills:1 key:ef7f2e2846
## Important Issues in Experiment Design These Modules Ignore
*** =video_link
//player.vimeo.com/video/198212060


--- type:VideoExercise lang:r aspect_ratio:62.5 xp:50 skills:1 key:5f5a6d0023
## Common Issues in Experiments
*** =video_link
//player.vimeo.com/video/199856354


--- type:MultipleChoiceExercise lang:r xp:50 skills:1 key:e302f4bddd
## Multiple Choice: Drawing conclusions from experiments
The transportation network company, Unter Technologies, is interested in increasing their revenue. They hypothesize that they would generate substantially greater profits if they lowered their costs and profit margins per each ride. 

Unter conducts an experiment on how its sales are sensitive to price reductions by offering a 25% discount for all of its services for one day. To their pleasant surprise, Unter sees a huge spike in sales and profits during the promotional day. 

Seeing the results of his experiment, Unter's CEO decides that Unter should lower its prices permanently. Why might this conclusion be a bit hasty?

*** =instructions
- Unter did not survey its customers to determine if they were enticed by Unter's lower prices.
- Unter does not have enough data to test the statistical significance of its findings.
- The treatment in Unter's experiment might not have the same effect as the treatment in their policy proposal.
- Unter's experiment didn't address the strategic responses of its competitors.
*** =sct
```{r}
msg1 = "Doing additional research to learn The Why behind a behavioral change can be helpful, but there's an experiment design issue to address first. Try again"
msg2 = "This question doesn't say how much data Unter used, but there's another experiment design issue to consider, so try again"
msg3 = "Correct! Unter's experiment specifically tests the effect on profits of a temporary price reduction in its services, rather than the effect of a permanent price reduction. Profits increased during the promotional period, but there is no evidence that profits would remain higher over a longer period of time."
msg4 = "While Unter's profits are a function of its competitors' behavior, there is another experiment design issue that is more appropriate. Try again"
test_mc(correct = 3, feedback_msgs = c(msg1,msg2,msg3,msg4))
```

--- type:MultipleChoiceExercise lang:r xp:50 skills:1 key:fa050e19b1
## Multiple Choice: Generalizing experimental results
Unter's CEO is still convinced that lowering the prices of their services might increase the company's profits. He is hesitant to lower Unter's prices for a longer period, since the company would risk losing a great deal of income if the price-drop did not substantially increase revenue. 

Therefore, to understand how a long-term price drop would affect Unter's profits, the CEO of Unter decides to drop the prices of their services in one city. Over the course of a year, Unter's profits in that city rose substantially. Unter's CEO is now convinced that he should drop Unter's prices nation-wide. Why might this conclusion still be too hasty?

*** =instructions
- Unter's experiment did not pay attention to how its changes affected its competitors.
- The context in Unter's experiment might not generalize to other contexts affected by Unter's policy proposal.
- Unter's experiment did not specify why people might be enticed by lower prices.
- Unter's experiment did not control for noncompliance.
*** =sct
```{r}
msg1 = "The company probably does exist in a complex market, but there's something else to worry about without that consideration. Try again"
msg2 = "Correct! Unter's experimental results may have been specific to the city that Unter conducted the experiment in. People from other cities may not be as sensitive to price-reductions as the city in Unter's experiment."
msg3 = "Doing some qualitative research to learn The Why behind a behavior change can be helpful, but there's an experiment design issue to address first. Try again"
msg4 = "The question assumes the CEO can successfully set prices in his own company, so if that's true, what other answer might still be a design issue we need to worry about?"
test_mc(correct = 2, feedback_msgs = c(msg1,msg2,msg3,msg4))
```

--- type:VideoExercise lang:r aspect_ratio:62.5 xp:50 skills:1 key:0f5e7a1e15
## Difficulties in Performing Randomized Experiments
*** =video_link
//player.vimeo.com/video/200052274


--- type:MultipleChoiceExercise lang:r xp:50 skills:1 key:7c1f83a32b
## Multiple Choice: Ethics in experiments
Which of the following causal questions is possible to answer with a randomized experiment (without major ethical concerns)?

*** =instructions
- The effect of losing one's job on mental health.
- The effect of religion on financial well-being.
- The effect of social isolation on academic performance.
- The effect of financial incentives on athletic performance.
*** =sct
```{r}
msg1 = "This would definitely cause major ethical concerns because randomly assigning people to lose their job is not allowed.Try again"
msg2 = "This is a very personal topic to many people, and likely to involve big ethical concerns, so try again"
msg3 = "Randomly assigning people to social isolation is incredibly unethical, so look for a better answer!"
msg4 = "Correct! Each of the other experiments is not only difficult to create, but could cause substantial harm to an experiment's participants."
test_mc(correct = 4, feedback_msgs = c(msg1,msg2,msg3,msg4))
```

--- type:MultipleChoiceExercise lang:r xp:50 skills:1 key:9ca8c3a544
## Common Issues in Interpretation I: Box-and-whisker plots
Not only are there common issues in experiments, but there are many common ways that people misinterpret statistical results. Box-and-whisker plots (also known as box plots) are a great tool for summarizing data, but can be a source of confusion for new analysts. To refamiliarize yourself with boxplots, You can read our summary of them in the Chapter 5 exercise, "Practice working with eGulf," or on various websites, such as Wikipedia (https://en.wikipedia.org/wiki/Box_plot).

Following an experiment regarding how computer keyboard layouts affect people's typing speed, the principal investigator, Kurt Devorak, compared the hand sizes among respondents in the treatment groups and control groups to make sure his sample was balanced. He plotted the results in two box-and-whiskers plots (illustrated in the R workspace). Which of the following conclusions **cannot** be determined from these boxplots?

*** =instructions
- The maximum hand size in the treatment group is identical to the maximum hand size in the control group.
- The third quartile hand size in the treatment group is identical to the third quartile hand size in the control group.
- The median hand size in the treatment group is identical to the median hand size in the control group.
- The mean hand size in the treatment group is identical to the mean hand size in the control group.

*** =pre_exercise_code
```{r}
set.seed(1)
rtnorm <- function(n, mean, sd, min = -Inf, max = Inf){
  qnorm(runif(n, pnorm(min, mean, sd), pnorm(max, mean, sd)), mean, sd)
}
options(scipen=1)
n=200
Control<-round(rtnorm(n,7.44,sd=.5,min = 6,max=9),2)
Treatment<-Control
#increase values between median and third quartile
  Treatment<-ifelse(Treatment<quantile(Treatment,.74)& Treatment>quantile(Treatment,.51),quantile(Treatment,.74),Treatment)
#increase values between median and first quartile
  Treatment<-ifelse(Treatment<quantile(Treatment,.49)& Treatment>quantile(Treatment,.26),quantile(Treatment,.49),Treatment)
  boxplot(Treatment,Control,names=c("Treatment","Control"),main="Hand Sizes (inches) across Experimental Groups")
```

*** =sample_code
```{r}
  boxplot(Treatment,Control,names=c("Treatment","Control"),main="Hand Sizes (inches) across Experimental Groups")
```
*** =sct
```{r}
msg1 = "The maximum value in both plots is indicated by the dot above the whiskers. These are identical. Try again"
msg2 = "The third quartile in both plots is indicated by the top edge of the boxes. These are identical. Try again"
msg3 = "The median value in both plots is indicated by the thick line within the box. These are identical. Try again"
msg4 = "Correct! Box-and-whiskers plots do not show the mean values in a sample. Although these two plots have identical quartile distributions, medians, and ranges, the mean values are actually different (7.8 in the treatment and 7.5 in the control)."
test_mc(correct = 4, feedback_msgs = c(msg1,msg2,msg3,msg4))
```


--- type:NormalExercise lang:r aspect_ratio:62.5 xp:50 skills:1 key:008ec0b3c5
## Common Issues in Interpretation II: Non-normal distributions
Many statistical models assume that the data is normally distributed like in a bell curve, where a) most observations lie around the mean, b) the frequency of data that falls far from the mean is inversely related with its distance from the mean, and c) the distribution of observations above and below the mean are relatively identical. Often times, results from experiments are non-normal, which can lead to spurious findings. Therefore, data transformations often need to be made to account for such problems.

On the TV show, Kitchen Nightmares, Bordon Hannity tries to revive failing restaurants by offering advice to their management and chefs. To measure the effectiveness of Hannity's efforts, a fan measures the average reviews of one of Hannity's revived restaurants before and after he offers them advice. With these results:

*** =instructions
- Determine whether the reviews are normally distributed
- Determine the average treatment effect of Hannity's efforts on restaurant reviews

*** =pre_exercise_code
```{r}
Reviews_Before<-sample(c(rep(1,23),rep(2,34),rep(3,28),rep(4,6),rep(5,34)))

Reviews_After<-sample(c(rep(1,37),rep(2,4),rep(3,26),rep(4,44),rep(5,14)))
```
*** =sample_code
```{r}
#In previous exercises, we encouraged you to explore the data before running any treatment tests. With the current data, we know that respondents were allowed to leave reviews on a likert-scale from 1-5, where 1 signals that the reviewers strongly disliked the restaurant and 5 signals that reviewers strongly liked the restaurant. However, let's try to be bad statisticians and dive into the results before we explore the data. What do the means look like?
  mean(Reviews_Before)
  mean(Reviews_After)
#They're identical! Does this mean that Hannity's show had no effect on this restaurant's ratings? Maybe not, let's plot the kernel densities of the reviews in the treatment and control groups to see how they are distributed.
  plot(density(Reviews_Before))
  plot(density(Reviews_After))
  
Solution1<-" " #Do these these look approximately normally distributed? Answer with "Yes" or "No"

#There are many ways to deal with multimodal data. One strategy when data are multimodally distributed on a likert scale is to collapse the response options. If it is not important to differentiate the level of dislike for a restaurant, we can collapse our scale to measure whether respondents liked or disliked the restaurant at all. In this case, we would want to collapse response options 1 and 2 into a single response option to represent negative opinions, and 4 and 5 into a single response option to represent positive opinions. Let's create a new variable with three parameters, where 1= Negative (1 or 2), 2= Ambivalent (3), and 3= Positive (4 or 5)
  
#To do this for the Reviews_Before variable, we can use the ifelse function. The below syntax says, if Reviews_Before==1 or 2, set Reviews_Before to 1, otherwise just return Reviews_Before.
    Reviews_Before<-ifelse(Reviews_Before==1 | Reviews_Before==2,1,Reviews_Before) 
    Reviews_Before<-ifelse(Reviews_Before==3,2,Reviews_Before) 
    Reviews_Before<-ifelse(Reviews_Before==4 | Reviews_Before==5,3,Reviews_Before)
  
#Now use the above syntax as a guide to rescale the Reviews_After variable.
    
#Once you have rescaled the Reviews_After variable, run a t.test between the Reviews_After and Reviews_Before variables, and return the p.value (use t.test()$p.value)

Solution2<- #pvalue of t.test on newly rescaled variables
```
*** =solution
```{r}
  Solution1<-"No"

    Reviews_Before<-ifelse(Reviews_Before==1 | Reviews_Before==2,1,Reviews_Before) 
    Reviews_Before<-ifelse(Reviews_Before==3,2,Reviews_Before) 
    Reviews_Before<-ifelse(Reviews_Before==4 | Reviews_Before==5,3,Reviews_Before)
    Reviews_After<-ifelse(Reviews_After==1 | Reviews_After==2,1,Reviews_After) 
    Reviews_After<-ifelse(Reviews_After==3,2,Reviews_After) 
    Reviews_After<-ifelse(Reviews_After==4 | Reviews_After==5,3,Reviews_After)

  Solution2<-t.test(Reviews_After,Reviews_Before)$p.value
```
*** =sct
```{r}
test_object("Solution1")
test_object("Solution2")
success_msg("Good work! If we assume that the difference in respondent's degrees of positive or negative feelings towards the restaurant are non-meaningful, we can conclude that Hannity's restaurant revival is effective (i.e. it increases the number of postive versus negative reviews). If the degree of positive or negative feelings towards restaurants are meaninginful, then we do not have enough information to reject the null hypothesis (i.e. evidence does not support the claim that Hannity's contributions leads restaurants to get better reviews")
```