--- 
title       : "Experiments 5: Practice with Published Experiments - The Oregon Health Insurance Experiment"
description : "These videos and exercises offer practice in skills used in analyzing experimental data."

--- type:VideoExercise lang:r aspect_ratio:62.5 xp:50 skills:1 key:46fda452f0
## The Design of the Oregon Health Experiment.
*** =video_link
//player.vimeo.com/video/198212086

--- type:VideoExercise lang:r aspect_ratio:62.5 xp:50 skills:1 key:47f92415c5
## Reading Average Treatment Effect & Confidence Intervals in a Table: Depression in the Oregon Health Experiment
*** =video_link
//player.vimeo.com/video/198212094

--- type:MultipleChoiceExercise lang:r xp:50 skills:1 key:2e74d95e94
## Multiple Choice: Confidence Intervals
Why might it be risky to implement a policy that has zero in the confidence interval of its average treatment effect?

*** =instructions
- Because the confidence interval equation defaults to 0 with large amounts of data.
- Because having zero in one's confidence interval implies that a treatment effect could have a positive or negative effect on the outcome of interest.
- Because having zero in one's confidence interval implies that a treatment effect has no effect.
- It's incorrect to think this would be risky; as long as relationship between the policy and its treatment effect are within the desired range, a confidence interval that includes zero is not a problem.
*** =sct
```{r}
msg1 = "This is not true. With tremendous sample size (N > 1,000,000), confidence intervals rarely cross 0, even for relatively insignificant treatment effects. Try again."
msg2 = "Correct! It would be very risky to implement a policy that could potentially have the opposite effect than you intended."
msg3 = "This is often how such confidence intervals are interpreted, but this is a mistake. A confidence interval that contains zero does is not evidence that there is no treatment effect, but that it is uncertain whether there is a treatment effect. Try again."
msg4 = "A confidence interval that contains zero is inherently problematic.  Try again."
test_mc(correct = 2, feedback_msgs = c(msg1,msg2,msg3,msg4))
```

--- type:MultipleChoiceExercise lang:r xp:50 skills:1 key:533b160157
## Multiple Choice: Significance
What is the difference between a statistically significant effect and a clinically (or "substantively") significant effect?

*** =instructions
- Statistically significant effects are bigger than clinically significant effects.
- Statistically significant effects don't necessarily matter, whereas clinically significant effects do matter.
- Statistically significant effects have smaller confidence intervals than clinically significant effects.
- Statistically significant effects always matter, whereas clinically significant effects sometimes don't matter.
*** =sct
```{r}
msg1 = "This might be true in some individual cases, but it's not true in general, so try again"
msg2 = "Correct! Statistical significance indicates whether an effect is likely real (probably not observed just by chance), but it does not necessarily indicate whether it is meaningful. With enough data, a study might find a statistically significant effect of season on people's preference to buy tissue paper, but the effect might not be important (for example, if people were 0.2% more likely to buy tissue paper in the spring over the fall)."
msg3 = "This might be true in some individual cases, but it's not true in general, so try again."
msg4 = "This is not how you should interpret treatment effects, so look again!"
test_mc(correct = 2, feedback_msgs = c(msg1,msg2,msg3,msg4))
```

--- type:NormalExercise lang:r aspect_ratio:62.5 xp:50 skills:1 key:0fa656dc08
## Practice Reading Tables (Part I)
Now let's practice reading tables from the Oregon Health Experiment. As described in the previous video, the Oregon Health Experiment studies the effect of getting Medicaid coverage on respondent's health, so let's look at some mental health results in the data.

The dataframe "OregonHealthResults" indicates results from the Oregon Health Experiment (in percentages) pertaining to depression. To get a good sense of how this was affected in the experiment, let's review some of its results and calculate some of the numbers ourselves. 

One of the most interesting results of any experiment is the Average Treatment Effect relative to the mean value of a parameter. Therefore, let's use this dataframe to look at two basic items: The average positive screen value in the control group, and the Average Treatment Effect of Medicaid on positive screen values. These calculations will give you a deeper sense of the impact of health insurance on basic mental health.


*** =instructions
- What is the mean value of positive screening in the control group?
- What is the average treatment effect of medicaid coverage on positive screening? 
- Given the mean value of positive screening in the control group and the average treatment effect of medicaid coverage on positive screening, estimate the mean value of positive screening in the treatment group.

*** =hint
- You can calculate the mean value of positive screening in the treatment group by adding the mean value in the control group with the average treatment effect.

*** =pre_exercise_code
```{r}
OregonHealth<-data.frame(c("Blood pressure - Systolic (mm Hg)","Blood pressure - Diastolic (mm Hg)","Blood pressure - Elevated (%)","Hypertension - Diagnosis after lottery (%)","Hypertension - Current use of medication for hypertension (%)","Cholesterol - Total level (mg/dl)","Cholesterol - High total level (%)","Cholesterol - HDL level (mg/dl)","Cholesterol - Low HDL level (%)","Hypercholesterolemia - Diagnosis after lottery (%)","Hypercholesterolemia - Current use of medication for high cholesterol level (%)","Glycated hemoglobin - Level (%)","Glycated hemoglobin - Level ???6.5% (%)","Diabetes - Diagnosis after lottery (%)","Diabetes - Current use of medication for diabetes (%)","Depression - Positive screening result (%)","Depression - Diagnosis after lottery (%)","Depression - Current use of medication for depression (%)","Framingham risk score - Overall","Framingham risk score - High-risk diagnosis","Framingham risk score - Age 50-65 yr"),
              c(119.3,76,16.3,5.6,13.9,204.1,14.1,47.6,28,6.1,8.5,5.3,5.1,1.1,6.4,30,4.8,16.8,8.2,11.6,13.9),
              c(-0.52,-0.81,-1.33,1.76,0.66,2.2,-2.43,0.83,-2.82,2.39,3.8,0.01,-0.93,3.83,5.43,-9.15,3.81,5.49,-0.21,1.63,-0.37),
              c(-2.97,-2.65,-7.16,-1.89,-4.48,-3.44,-7.75,-1.31,-10.28,-1.52,-0.75,-0.09,-4.44,1.93,1.39,-16.7,0.15,-0.46,-1.56,-1.11,-2.64),
              c(1.93,1.04,4.49,5.4,5.8,7.84,2.89,2.98,4.64,6.29,8.35,0.11,2.59,5.73,9.48,-1.6,7.46,11.45,1.15,4.37,1.9),
              c(0.68,0.39,0.65,0.34,0.8,0.45,0.37,0.45,0.46,0.23,0.1,0.82,0.61,0.001,0.008,0.02,0.04,0.07,0.76,0.24,0.75))
names(OregonHealth)<-c("Variable","Mean_Value_in_Control_Group","Change_with_Medicaid_Coverage","95%_Confidence_Interval_(lower_bound)","95%_Confidence_Interval_(Upper_Bound)","P_Value")
OregonHealthResults<-OregonHealth[16:18,]
OregonHealthResults$Variable<-c("Positive_Screening","Diagnosis_After_Lottery","Current_Use_Of_Medication")
rm(OregonHealth)
```
*** =sample_code
```{r}
# Let's being by printing the results from the OregonHealthResults dataframe
    print(OregonHealthResults) 

# Now let's practice reading the table to identify the following:

# Question 1: Write the mean value of positive screening in the control group (do not round).
#---- Question 1-------------------------------------#
    Solution1<-
#----------------------------------------------------#

# Question 2: Write the average treatment effect of medicaid coverage on positive screening.
#---- Question 2-------------------------------------#
    Solution2<- 
#----------------------------------------------------#

# Question 3: Use solution1 and Solution2 to calculate the mean value of positive screening in the treatment group.
#---- Question 3-------------------------------------#
Solution3<-  
#----------------------------------------------------#

```

*** =solution
```{r}
Solution1<-OregonHealthResults[OregonHealthResults$Variable=="Positive_Screening",]$Mean_Value_in_Control_Group
Solution2<-OregonHealthResults[OregonHealthResults$Variable=="Positive_Screening",]$Change_with_Medicaid_Coverage
Solution3<-Solution1+Solution2

```

*** =sct
```{r}
test_object("Solution1")
test_object("Solution2")
test_object("Solution3")
success_msg("Good work!")
```

--- type:NormalExercise lang:r aspect_ratio:62.5 xp:50 skills:1 key:5fa7d35f25
## Practice Reading Tables (Part II)
Now that we know how an Average Treatment Effect is related to mean values in the treatment and control groups, we should examine the reliability of these results. Therefore, let's examine the 95% confidence interval lower bound, and statistical significance for one of the table's Average Treatment Effects.

*** =instructions
- What is the the 95% confidence interval lower bound for the treatment effect of Medicaid coverage on current use of medication?
- Which average treatment effect of Medicaid coverage on depression was not statistically significant? You can write this out as a line of text.

*** =pre_exercise_code
```{r}
OregonHealth<-data.frame(c("Blood pressure - Systolic (mm Hg)","Blood pressure - Diastolic (mm Hg)","Blood pressure - Elevated (%)","Hypertension - Diagnosis after lottery (%)","Hypertension - Current use of medication for hypertension (%)","Cholesterol - Total level (mg/dl)","Cholesterol - High total level (%)","Cholesterol - HDL level (mg/dl)","Cholesterol - Low HDL level (%)","Hypercholesterolemia - Diagnosis after lottery (%)","Hypercholesterolemia - Current use of medication for high cholesterol level (%)","Glycated hemoglobin - Level (%)","Glycated hemoglobin - Level ???6.5% (%)","Diabetes - Diagnosis after lottery (%)","Diabetes - Current use of medication for diabetes (%)","Depression - Positive screening result (%)","Depression - Diagnosis after lottery (%)","Depression - Current use of medication for depression (%)","Framingham risk score - Overall","Framingham risk score - High-risk diagnosis","Framingham risk score - Age 50-65 yr"),
              c(119.3,76,16.3,5.6,13.9,204.1,14.1,47.6,28,6.1,8.5,5.3,5.1,1.1,6.4,30,4.8,16.8,8.2,11.6,13.9),
              c(-0.52,-0.81,-1.33,1.76,0.66,2.2,-2.43,0.83,-2.82,2.39,3.8,0.01,-0.93,3.83,5.43,-9.15,3.81,5.49,-0.21,1.63,-0.37),
              c(-2.97,-2.65,-7.16,-1.89,-4.48,-3.44,-7.75,-1.31,-10.28,-1.52,-0.75,-0.09,-4.44,1.93,1.39,-16.7,0.15,-0.46,-1.56,-1.11,-2.64),
              c(1.93,1.04,4.49,5.4,5.8,7.84,2.89,2.98,4.64,6.29,8.35,0.11,2.59,5.73,9.48,-1.6,7.46,11.45,1.15,4.37,1.9),
              c(0.68,0.39,0.65,0.34,0.8,0.45,0.37,0.45,0.46,0.23,0.1,0.82,0.61,0.001,0.008,0.02,0.04,0.07,0.76,0.24,0.75))
names(OregonHealth)<-c("Variable","Mean_Value_in_Control_Group","Change_with_Medicaid_Coverage","95%_Confidence_Interval_(lower_bound)","95%_Confidence_Interval_(Upper_Bound)","P_Value")
OregonHealthResults<-OregonHealth[16:18,]
OregonHealthResults$Variable<-c("Positive_Screening","Diagnosis_After_Lottery","Current_Use_Of_Medication")
rm(OregonHealth)
```
*** =sample_code
```{r}
# Let's begin by printing the results from the OregonHealthResults dataframe
    print(OregonHealthResults)

# Now let's practice reading the table and identify the following:

# Question 1: Write the 95% confidence interval lower bound for the treatment effect of Medicaid coverage on current use of medication.
#---- Question 1-------------------------------------#
    Solution1<-
#----------------------------------------------------#

# Question 2: Write the name of the variable whose average treatment effect was not statistically significant (must be a string).
#---- Question 2-------------------------------------#
    Solution2<-""
#----------------------------------------------------#
```

*** =solution
```{r}

Solution1<-OregonHealthResults[OregonHealthResults$Variable=="Current_Use_Of_Medication",]$`95%_Confidence_Interval_(lower_bound)`
Solution2<-OregonHealthResults[OregonHealthResults$P_Value>.05,]$Variable

```


*** =sct
```{r}
test_object("Solution1")
test_object("Solution2")
success_msg("Good work!")
```
--- type:VideoExercise lang:r aspect_ratio:62.5 xp:50 skills:1 key:743f5494ef
## A Note on Heteroskedasticity
*** =video_link
//player.vimeo.com/video/205124690


--- type:MultipleChoiceExercise lang:r xp:50 skills:1 key:92568b1f93
## Practice identifying heteroskedastic outcomes
An easy way to identify whether the variability of a dependent variable is unequal across the range of values in an independent variable is to examine these relationships in a scatter plot. The R Workspace contains plots for income's relationship with three different predictors: Education, Age, and City Population. You can see each plot by clicking the "Next Plot" and "Previous Plot" buttons below the plot. Within the plots, each dot represents an observation from our sample. Using these plots, determine the independant variable with which income is most heteroskedastic.

*** =instructions
- Education
- Age
- City Population

*** =pre_exercise_code
```{r}
set.seed(1)
mround <- function(x,base){ 
  base*round(x/base) 
} 
rtnorm <- function(n, mean, sd, min = -Inf, max = Inf){
  qnorm(runif(n, pnorm(min, mean, sd), pnorm(max, mean, sd)), mean, sd)
}
options(scipen=1)
n=900
Sample<-as.data.frame(matrix(nrow=n,ncol=4))
names(Sample)<-c("Income","Education","Age","City_Population")
Sample$Education<-rtnorm(n, mean=15, sd=3, min = 10, max = 20)
Sample$Age<-round(rtnorm(n, mean=45, sd=10, min = 15, max = 92),0)
Sample$City_Population<-rtnorm(n, mean=10000, sd=1000, min = 8000, max = 15000)
Sample$Income<-(20000*Sample$Education+20*Sample$City_Population+Sample$Age*5000+ifelse(Sample$Age>45,rtnorm(n, mean=1000, sd=1000*(45-Sample$Age)^2, min = -500000, max = 1000000),0))/2
  
plot(Sample$Age,Sample$Income)
plot(Sample$Education,Sample$Income)
plot(Sample$City_Population,Sample$Income)

```

*** =sample_code
```{r}
#Relationship between Age and Income
  plot(Sample$Age,Sample$Income)

#Relationship between Education and Income
  plot(Sample$Education,Sample$Income)

#Relationship between City Population and Income
  plot(Sample$City_Population,Sample$Income)
```
*** =sct
```{r}
msg1 = "Almost, try again."
msg2 = "Good job! This is an example of a heteroskedastic treatment effect. Among our sample, income was slightly correlated with age, but only varied slightly until respondents reached their late 50s. If we were to ignore this trend while studying the relationship between age and income, it could confound or bias our results."
msg3 = "Almost, try again."
test_mc(correct = 2, feedback_msgs = c(msg1,msg2,msg3))
```




--- type:VideoExercise lang:r aspect_ratio:62.5 xp:50 skills:1 key:6a474387e6
## Reading Average Treatment Effect & Confidence Intervals in a Table: Cholesterol in the Oregon Health Experiment
*** =video_link
//player.vimeo.com/video/198212098


--- type:MultipleChoiceExercise lang:r xp:50 skills:1 key:5765daa781
## Multiple Choice: Interpreting effect sizes
The head of marketing at Bizer, Inc. is trying to decide whether to include a new cholesterol medicine's average treatment effect in the company's advertisements for that medicine. He knows that on average, the medicine reduces cholesterol in respondents by about 1mg/dL and knows that the effect is statistically different from zero. However, he's not sure whether that number is large or small; that is, whether the medicine's average treatment effect is worth advertising. 

If cholesterol is highly variable in the population (about 70% of the population has cholesterol between 160mg/dL and 240mg/dL), does an average treatment effect of 1mg/dL seem relatively small or relatively large? In other words, when an average treatment effect is substantially smaller than the standard error of the mean for a parameter, how would we typically interpret the size of that average treatment effect?


*** =instructions
- The average treatment effect is large.
- The average treatment effect is small.
*** =sct
```{r}
msg1 = "Although a 1mg/dL reduction in cholesterol is statistically significant and might be substantial for some members of the population, it is not a very large reduction relative to the amount of variance in cholesterol in the population. Try again."
msg2 = "Correct! When an average treatment effect is small relative to the standard error of a sample mean, it suggests that the average treatment effect is small, relative to the amount of variance typically found in that parameter. This thought process is especially helpful when deciding if an estimated effect is clinically significant (even if it is statistically significant)."
test_mc(correct = 2, feedback_msgs = c(msg1,msg2))
```

--- type:NormalExercise lang:r aspect_ratio:62.5 xp:50 skills:1 key:82150d848b
## Practice working with the Oregon Health Experiment
Let's say you're wondering how much mens' and womens' health differed in the experiment, and what the data says about Medicaid's effects on those differences, so you decide to look at some of the basic health numbers in the data. A simulated version of the Oregon Health Insurance Experiment data, `OHIE`, is available in the workspace. With this dataframe, make some quick calculations to learn about any differences that might be linked to gender by doing the following:

*** =instructions
- Test if there is a statistically significant difference in gender between the treament and control groups (i.e. whether gender is balanced between the treatment and control group). 
- Determine if the treatment and control groups have significantly different average values for the following health outcomes: systolic blood pressure (variable `bp_sar_inp`), and diastolic blood pressure (variable `bp_dar_inp`).

*** =hint
- For the first question, you will need to `subset` the `OHIE` data frame by treatment. 
- It may help to create a new dataframe called `TreatmentGroup` that contains only those selected to be able to receive Medicaid in the treatment group, and a new data frame called `ControlGroup` that contains only those not selected for Medicaid in the control group.
- For each of the following questions, use the `t.test` function where (mu=0) to statistically test whether the treatment and control group are different from each other.


*** =pre_exercise_code
```{r}
set.seed(1)
load(url('http://s3.amazonaws.com/assets.datacamp.com/production/course_1566/datasets/OHIEexperimental.Rda'))
OHIE <- OHIE[!is.na(OHIE$treatment),c("id","treatment","gender_inp","bp_sar_inp","bp_dar_inp")]

```

*** =sample_code
```{r}
# Since the OHIE dataset is fairly big, let's start to get a sense of it by looking at its structure
    str(OHIE)

# Since we are going to run several t-tests, let's make the syntax below simpler by separating the OHIE dataset into treatment and control groups.
    TreatmentGroup <- OHIE[OHIE$treatment==1, ]
    ControlGroup <- OHIE[OHIE$treatment==0, ]

# Since we are still early in our unit, we will help you answer the first question. We need to determine whether gender is balanced between the treatment and control groups. This requires a two sample t-test.
#---- Question 1a------------------------------------#
    Solution1a<-t.test(TreatmentGroup$gender_inp,ControlGroup$gender_inp,mu=0)
#----------------------------------------------------#

# To determine whether the t-test shows that gender is balanced, we need to determine whether its results are statistically significant. If the 95% confidence interval includes 0, the results are not statistically significant (i.e. we cannot be confident that there is a difference between the treatment and control group). Based on the confidence interval in Solution 1a, do the results appear statistically significant? Answer with "Yes" or "No." 
#---- Question 1b------------------------------------#
    Solution1b<-""  
#----------------------------------------------------#

# A more direct way to test for statistical significance is to just look at the p-value of the t-test. To directly compute the p-value for solution 1a, we could have just typed: "Solution1a <- t.test(TreatmentGroup$gender_inp, ControlGroup$gender_inp, mu=0)$p.value". For Question 2, use a t-test to determine whether the treatment had a statistically significant effect on systolic blood pressure, and return the p-value.
#---- Question 2a------------------------------------#
    Solution2a<-  
#----------------------------------------------------#

# A p.value of less than .05 is generally considered statistically significant. Based on your results in Solution 2a, did the experiment yield a statistically signficant average treatment effect on systolic blood pressure? Answer with "Yes" or "No."
#---- Question 2b------------------------------------#
    Solution2b<-""  
#----------------------------------------------------#

# Question 3a: Return the p-value of a t-test that measures the average treatment effect on diastolic blood pressure.
#---- Question 3a------------------------------------#
    Solution3a<-   
#----------------------------------------------------#

# Question 3b: Based on your t-test, does the treatment have a statistically significant effect on diastolic blood pressure? Answer with "Yes" or "No."
#---- Question 3b------------------------------------#
    Solution3b<-""   
#----------------------------------------------------#


```

*** =solution
```{r}
TreatmentGroup <- OHIE[OHIE$treatment==1, ]
ControlGroup <- OHIE[OHIE$treatment==0, ]
Solution1a<- t.test(TreatmentGroup$gender_inp,ControlGroup$gender_inp,mu=0)
Solution1b<-"Yes"
Solution2a<- t.test(TreatmentGroup$bp_sar_inp,ControlGroup$bp_sar_inp,mu=0)$p.value
Solution2b<-"No"
Solution3a<- t.test(TreatmentGroup$bp_dar_inp,ControlGroup$bp_dar_inp,mu=0)$p.value
Solution3b<-"Yes"

```

*** =sct
```{r}
test_object("Solution1a")
test_object("Solution1b")
test_object("Solution2a")
test_object("Solution2b")
test_object("Solution3a")
test_object("Solution3b")

success_msg("Good work!")
```



--- type:NormalExercise lang:r aspect_ratio:62.5 xp:50 skills:1 key:4f3ebb92a2
## Practice working with eGulf
To increase their commissions, the popular online auctioneer, eGulf, wants to help its sellers increase the final sales prices of their merchandise. For each auction on eGulf, sellers are allowed to post up to 10 pictures of an item that they are selling. Past studies suggest that customers' appraisals of merchandise (and subsequently, final bids of merchandise) improve when sellers include more pictures of their merchandise; therefore, eGulf considers raising the limit on the number of pictures that a seller can post on an auction's webpage.

To test whether eGulf should allow sellers to post more than 10 pictures of their items, eGulf conducts an experiment: eGulf finds a random sample of dedicated used WePhone sellers who typically post 10 pictures of their used WePhones, and temporarily allows them to post up to 15 pictures for each auction of their WePhones. Using the results of this experiment, listed in dataset `eGulf`, test whether there is an added benefit to posting more than 10 pictures on WePhone auctions. Specifically:

*** =instructions
- Determine the difference in means betwen the `Final_Sales_Price` of sellers who posted more than 10 pictures (`Seller_Opted_Into_Treatment`==1) on their auctions, and those who had not.
- Statistically test whether sellers who posted more than 10 pictures (`Seller_Opted_Into_Treatment`==1) on their WePhone auctions (`Final_Sales_Price`) sold their WePhones at higher prices than sellers who posted 10 or fewer pictures on their WePhone auctions (`Seller_Opted_Into_Treatment`==0)

*** =hint
- For the first question, you will need to `subset` the `eGulf` data frame by treatment. 
- It may help to create a new dataframe called `TreatmentGroup` that contains only sellers who posted 11 or more pictures in the treatment group, and a new data frame called `ControlGroup` that contains only sellers who posted fewer than 11 pictures in the control group.
- Use the `t.test` function and report the p-value to statistically test whether the treatment and control group yielded different sales prices.

*** =pre_exercise_code
```{r}
# Initialize dataframe
set.seed(1)
n <- 1000
eGulf <- as.data.frame(matrix(0, ncol=5,nrow=n))
colnames(eGulf) <- c("Seller_Offered_Treatment","Seller_Opted_Into_Treatment","Seller_Feedback_Count","Seller_Feedback_Score","Auction_Final_Sales_Price")
# Simulate baseline data
  # Make seller feedback count
    eGulf$Seller_Feedback_Count<-as.integer(abs(round(rnorm(n,200,100))))
  # Make feedback score loosely correlated with feedback count 
    eGulf$Seller_Feedback_Score<-round(.9+.1*eGulf$Seller_Feedback_Count/max(eGulf$Seller_Feedback_Count)-rbeta(100,1,10),2)
  # Make seller offered treatment loosely correlated with with feedback count 
    eGulf$Seller_Offered_Treatment<-rbinom(n,1,.5+eGulf$Seller_Feedback_Count/max(eGulf$Seller_Feedback_Count)/5)
  # Make seller opted_into_treatment correlated with seller_feedback_score
    eGulf$Seller_Opted_Into_Treatment<-ifelse(eGulf$Seller_Offered_Treatment==0,0,rbinom(n,1,eGulf$Seller_Feedback_Score^5)) #
  # Make final sales price also correlated with seller_feedback_score
    eGulf$Final_Sales_Price<-round(rlnorm(n, meanlog=log(400), sdlog = log(1.2))*eGulf$Seller_Feedback_Score)
```
*** =sample_code
```{r}
# To begin, let's examine the structure of the dataframe with the str function. 
    str(eGulf)

# To get an even better sense of the data, let's visualize the data graphically. Boxplots are often ideal for studying data in controlled experiments, as they offer a convenient means to compare two groups (the treatment and control). The following boxplots show the relationship between opting into treatment and final sales prices. The boxplot on the left represents data from the control group, whereas the boxplot on the right represents data from the treatment group. Within these boxplots, the bold middle line represents the median of the values within each group. The two lines above and below the median that complete the "box" represent the upper and lower quartiles of data in the dataset. The dots and "whiskers" at the edge of the plot represent the most extreme datapoints in the two datasets.
    boxplot(eGulf$Final_Sales_Price~eGulf$Seller_Opted_Into_Treatment)

# From the boxplot, it appears that sellers who opted into treatment (the box on the right), sold their WePhones at slightly higher prices. However, let's make sure of this by computing the mean differences. Subtract the mean sales prices of the control group from the mean sales prices in the treatment group. You may want to subset the treatment and control groups into separate dataframes before completing this next step.
#---- Question 1-------------------------------------#
    Solution1<-   
#----------------------------------------------------#

# If you answered Solution1 correctly, it should appear that sellers in the treatment group sold their Wephones at higher prices than sellers in the control group. However, let's conduct a statistical test to make sure this finding is robust. Report the p-value of a t-test that tests whether opting into treatment (i.e. posting more than 10 pictures) increased final sales prices  
#---- Question 2-------------------------------------#
    Solution2<-   
#----------------------------------------------------#
  
 
```
*** =solution
```{r}
Solution1<-mean(eGulf$Final_Sales_Price[eGulf$Seller_Opted_Into_Treatment==1])-mean(eGulf$Final_Sales_Price[eGulf$Seller_Opted_Into_Treatment==0])
Solution2<-t.test(eGulf$Final_Sales_Price[eGulf$Seller_Opted_Into_Treatment==1],eGulf$Final_Sales_Price[eGulf$Seller_Opted_Into_Treatment==0])$p.value

```
*** =sct
```{r}
test_object("Solution1")
test_object("Solution2")
success_msg("Good work! Since the p-value is well below .05, it seems that our finding is robust; likely posting more photos was positively associated with final WePhone sales prices.")
```
--- type:VideoExercise lang:r aspect_ratio:62.5 xp:50 skills:1 key:ad358b17bb
## Where the Chapter Will Go From Here
*** =video_link
//player.vimeo.com/video/205124618
