--- 
title       : "Experiments 11: Putting it All Together"
description : "These Exercises Will Wrap up What we have learned about experiments"
  
--- type:VideoExercise lang:r aspect_ratio:62.5 xp:50 skills:1 key:a39c1976d4
## Bentley Video...
*** =video_link
//player.vimeo.com/video/
 

--- type:NormalExercise lang:r aspect_ratio:62.5 xp:100 skills:2 key:5b941b8026
## Putting it All Together
Often times, our experiments don't go as planned. This question will test what you have learned to resolve a question with realistic problems.

KittyCatch is a location-based augmented reality game developed by Meowtec. In this game, players use their cell phones' GPS abilities to locate and capture virtual kittens. Although KittyCatch is free to play, Meowtec generates revenue from KittyCatch through pop-up ads, that appear about once for every half mile the player walks while running the game, up to 5 times a day. Since Meowtec's revenue is highly dependent on how far people walk while playing KittyCatch, their primary goal for future updates to KittyCatch is to incentivize play. 

One strategy to increase user's play time is to increase the distance between KittyCatch's "points of kinterest;" that is, increase the distance that users must travel to capture new kittens. Many players stop their sessions of KittyCatch before they catch a single kitten, and may be turned off by the need to travel farther distances to catch a kitten, but Meowtec believes that the loss of play time for these players would be offset by the many players who consistently travel as far as needed to catch a kitten. 

To test this hypothesis, Meowtec constructs an experiment: For a small random sample of players in Springfield, Massachusetts, Meowtec increases the distance that players must walk to reach a point of kinterest (Specifically, from .25 miles to .5 miles). They compare these players to an equal random sample of players in Springfield, Massachusetts that have the default distance between points of kinterest (.25 miles). However, as they explored their data, they found a few problems with their experiment. With the dataset, `KittyCatch`, and the annotations in the R workspace, help Meowtec determine a valid and reliable average treatment effect from their experiment.

*** =instructions
- Balance the data
- Determine the average treatment effect

*** =pre_exercise_code
```{r}
set.seed(1)
#Create rnorm function that allows for random distribution with min and max
  rtnorm <- function(n, mean, sd, min = -Inf, max = Inf){
    qnorm(runif(n, pnorm(min, mean, sd), pnorm(max, mean, sd)), mean, sd)
  }
#Create rounding function that allows to round to numbers above 1
  mround <- function(x,base){ 
    base*round(x/base) 
  } 

n=688
#Create dataframe
  KittyCatch<-as.data.frame(matrix(nrow=n,ncol=4))
  names(KittyCatch)<-c("DistanceWalked","Age","TotalHoursPlayed","Treatment")
#Assign Treatment randomly
  KittyCatch$Treatment<-rbinom(n,1,.35)
#Assign Values to variables
 #DistanceWalked
  #Create Variable
    KittyCatch$DistanceWalked<-round(rtnorm(n=n, mean=.62, sd=.3, min = 0, max = 1.3),2)
  #Assign 4 Outliers to DistanceWalked among control
    KittyCatch$DistanceWalked[KittyCatch$Treatment==0][sample(sum(KittyCatch$Treatment==0),4)]<-round(rtnorm(n=4, mean=7.00, sd=.6, min = 5, max = 9),2)
  #Age
    #Create Variable
      KittyCatch$Age<-round(rtnorm(n=n, mean=14, sd=4, min = 10, max = 40))
    #Assign age to be correlated with DistanceWalked
      KittyCatch$Age<-round(KittyCatch$Age+(ifelse(KittyCatch$DistanceWalked>median(KittyCatch$DistanceWalked),2,-2)))
    #Assign age to be correlated with treatment above age 18
      KittyCatch$Age<-round(KittyCatch$Age+(ifelse(KittyCatch$Age>17 & KittyCatch$Treatment==1,4,0)))
    #Top code in control group
      KittyCatch$Age[KittyCatch$Treatment==0]<-ifelse(KittyCatch$Age[KittyCatch$Treatment==0]>18,18,KittyCatch$Age[KittyCatch$Treatment==0])
  #TotalHoursPlayed
    #Create Variable
      KittyCatch$TotalHoursPlayed<-round(rtnorm(n=n, mean=30, sd=10, min = 1, max = 50))
    #Make it correlated with distance walked
      KittyCatch$TotalHoursPlayed<-round(.8*KittyCatch$TotalHoursPlayed + .2*KittyCatch$TotalHoursPlayed*KittyCatch$DistanceWalked)
    #Randomly assign missing values, make the proportion larger in the treatment; and make the likelihood for NA correlate with larger values
      KittyCatch$TotalHoursPlayed[KittyCatch$Treatment==0]<-ifelse(rbinom(sum(KittyCatch$Treatment==0),1,(.05+.05*KittyCatch$TotalHoursPlayed[KittyCatch$Treatment==0]/mean(KittyCatch$TotalHoursPlayed))),NA,KittyCatch$TotalHoursPlayed[KittyCatch$Treatment==0])
      KittyCatch$TotalHoursPlayed[KittyCatch$Treatment==1]<-ifelse(rbinom(sum(KittyCatch$Treatment==1),1,(.25+.05*KittyCatch$TotalHoursPlayed[KittyCatch$Treatment==1]/mean(KittyCatch$TotalHoursPlayed,na.rm = T))),NA,KittyCatch$TotalHoursPlayed[KittyCatch$Treatment==1])
  #Make distance walked negatively correlated with treatment prior to balance, and positvely correlated with treatment after balance
      removed<-ifelse(KittyCatch$Age>17 | is.na(KittyCatch$TotalHoursPlayed),1,0)
      KittyCatch$DistanceWalked<-round(ifelse(removed==1 & KittyCatch$Treatment==0,1.5*KittyCatch$DistanceWalked,ifelse(removed==0 & KittyCatch$Treatment==1,1.3*KittyCatch$DistanceWalked,KittyCatch$DistanceWalked)),2)

```
*** =sample_code
```{r}
#Let's first take a naive estimate of the average treatment effect, before testing or correcting for balance in the dataset
  t.test(KittyCatch$DistanceWalked[KittyCatch$Treatment==1],KittyCatch$DistanceWalked[KittyCatch$Treatment==0])
  mean(KittyCatch$DistanceWalked[KittyCatch$Treatment==1])- mean(KittyCatch$DistanceWalked[KittyCatch$Treatment==0])
#It seems that there is substantial negative treatment effect; respondents walked less in the treatment than in the control. But is this finding reliable?

#Let's run the structure command to get a sense of the dataset. DistanceWalked is our outcome of interest, and Treatment is the key independent variable that we care about. TotalHoursPlayed refers to how many hours the user has put into KittyCatch in total.
  str(KittyCatch)
#Now let's examine our outcome of interest, DistanceWalked
  summary(KittyCatch$DistanceWalked)
#Notice that the maximum value is much larger than the mean, median, and 3rd quartile of the dataset. This outlier might bias our t.test results by violating its assumption that the data is normally distributed. See density plot below to observe the long right tail in the distribution of DistanceWalked.
  plot(density(KittyCatch$DistanceWalked))
#Since the dataset is relatively small, let's use the `head` and `order` commands to examine the highest DistanceWalked observations in the dataset
  head(KittyCatch[order(KittyCatch$DistanceWalked,decreasing=T),]) #This syntax shows us 6 observations, sorted by the largest value for DistanceWalked
#Notice that the four largest values for DistanceWalked are more than three times as large as the 5th largest value for DistanceWalked. These are clearly outliers. However, we should also note that all of these values are in the control group, so removing them from our dataset, as we had done with previous outliers, might also bias our estimates by lowering the mean values for the treatment group. Instead, let's top code these values so that they are still the largest values in the dataset, but so that they are not outliers and do not violate our assumptions of normality. Top coding can be arbitrary (and there are better methods for handling outliers), but for this assignment, let's just assign all DistanceWalked observations above a value of 2 to 2

#---- Question 1: Top code DistanceWalked to 2 ----#
   KittyCatch$DistanceWalked<- 
#----------------------------------------------------#
    
#If you completed question 1 correctly, the distribution of DistanceWalked should look much more normal
  plot(density(KittyCatch$DistanceWalked))
#We'll also want to make sure that our control groups and treatment groups are balanced. Let's compare the age of the control group and treatment group with boxplots
  boxplot(KittyCatch$Age[KittyCatch$Treatment==0],KittyCatch$Age[KittyCatch$Treatment==1],names=c("Control","Treatment"))
#Notice that the medians and interquartile ranges are very similar between the treatment and control groups, but that the largest values in the treatment group rest much higher than in the control group. It turns out that this did not happen by chance; by default, KittyCatch only requires respondents to enter their ages if they state that they are younger than 18; however, for the experiment, all respondents in the treatment group were required to enter their ages. It could be problematic if the ages among the treatment group are different than those in the control group, especially if Age is correlated with DistanceWalked, which, as demonstrated below, it is.
  cor(KittyCatch$Age,KittyCatch$DistanceWalked)
#This also seems to cause the treatment and control groups to be imbalanced with regard to age.
  t.test(KittyCatch$Age[KittyCatch$Treatment==0],KittyCatch$Age[KittyCatch$Treatment==1])
#There are two options to deal with this problem. One option is to ignore it, as we know that age was top coded in the control group, and may entirely be the cause of the imbalance between the treatment and control group. However, what if people who opted into the treatment group tended to be older on average? This could bias our results. A simple way to deal with this problem is to subset the data to only use observations we can be sure are similarly aged. Since we can only be sure of the age of respondents in the treatment group for those younger than 18, let's subset the dataset to only include respondents who are 17 or younger.
  
#---- Question 2: Subset KittyCatch to respondents who are 17 or younger ----#
  KittyCatch<-
#----------------------------------------------------------------------------#

#If you completed question 2 correctly, Age among the treatment and control groups should now be balanced
  t.test(KittyCatch$Age[KittyCatch$Treatment==0],KittyCatch$Age[KittyCatch$Treatment==1])
#Now, let's look for balance in the last variable we know about the treatment and control groups, TotalHoursPlayed.
  summary(KittyCatch$TotalHoursPlayed)
#Notice that 127 values are labeled "NA". These cases are missing information, in large part because KittyCatch allows users to opt out of recording how many hours they play. However, as shown below, TotalHours played appears to be correlated with DistanceWalked.
  cor(KittyCatch$TotalHoursPlayed,KittyCatch$DistanceWalked,use="complete")
#Moreover, NA values for TotalHoursPlayed appears at higher rates for respondents in the treatment group than in the control group
  mean(is.na(KittyCatch$TotalHoursPlayed[KittyCatch$Treatment==1]))
  mean(is.na(KittyCatch$TotalHoursPlayed[KittyCatch$Treatment==0]))
#While total hours played appears balanced among the treatment and control groups for those whose information is available (see t.test below), the fact that its missingness is correlated with DistanceWalked and patterned among the treatment and control groups is cause for concern. Specifically, this could suggest that respondents in the treatment group play more KittyCatch than those in the control group, and are likely to walk further with KittyCatch due to this. A simple solution to this problem then, is listwise deletion; that is, removing respondents from our sample who did not report their total hours played.
  t.test(KittyCatch$TotalHoursPlayed[KittyCatch$Treatment==1],KittyCatch$TotalHoursPlayed[KittyCatch$Treatment==0])

#---- Question 3: Subset KittyCatch to respondents who reported their total hours played ----#
  KittyCatch<-
#--------------------------------------------------------------------------------------------#

#Now that we have balanced our data on all known dimensions, let's return the average treatment effect, and test whether it is significant

#---- Question 4: Determine the average treatment effect (mean of treatment - mean of control) ----#
  Solution4<-
#--------------------------------------------------------------------------------------------------#

#---- Question 5: Determine the significance of the average treatment effect (via a t.test) -------#
  Solution5<-t.test()$p.value
#--------------------------------------------------------------------------------------------------#
  
```
*** =solution
```{r}
  KittyCatch$DistanceWalked<-ifelse(KittyCatch$DistanceWalked>2,2,KittyCatch$DistanceWalked)
  KittyCatch<-KittyCatch[KittyCatch$Age<=17,]
  KittyCatch<-KittyCatch[!is.na(KittyCatch$TotalHoursPlayed),]
  Solution4<-mean(KittyCatch$DistanceWalked[KittyCatch$Treatment==1])-mean(KittyCatch$DistanceWalked[KittyCatch$Treatment==0])
  Solution5<-t.test(KittyCatch$DistanceWalked[KittyCatch$Treatment==1],KittyCatch$DistanceWalked[KittyCatch$Treatment==0])$p.value
```
*** =sct
```{r}
test_object("Solution4")
test_object("Solution5")
success_msg("Good work! Our naive estimate of the average treatment effect was opposite of our estimate after we tried to balance the data. However, we had to ignore a large fraction of our observations to get to that conclusion. This is the difficulty of computing conditional average treatment effects with t.tests. However, this problem can be sidestepped with regression models. Our next module teaches how to use regression models in detail.")
```