--- 
title       : "Experiments 3: Randomized Experiments"
description : "These videos and exercises introduce randomized experiments."

--- type:VideoExercise lang:r aspect_ratio:62.5 xp:50 skills:1 key:9e9326a35b
## Randomized Experiments
*** =video_link
//player.vimeo.com/video/198212082


--- type:MultipleChoiceExercise lang:r xp:50 skills:1 key:243af9a90f
## Multiple Choice: Experimental design
Laurel is wondering if it would be more realistic to test phone durability through a bending test rather than a crushing test. To find out, she convinces four of her friends to join her at an electronics store and bend one of the two phones for a duration of 1 minute each. 

Two of her friends really want to try to bend the Universe S10, so Laurel assigns those two friends to bend the Universe S10, and her other two friends try bending the WePhone 10S. 

The Universe S10 was bent about 2 degrees more than the WePhone 10, so Laurel concludes that the average treatment effect of the WePhone 10S on bendability is -2 degrees. Of the following, which is the most problematic design error in this experiment?


*** =instructions
- Noncompliance
- Ethics
- Random assignment
- Not enough data

*** =sct
```{r}
msg1 = "Not in this case, as all her friends did what she asked. Try again."
msg2 = "There are probably ethical issues involved with knowingly breaking things in a store, but from a scientific perspective, there's a bigger experiment design issue than that. Try again."
msg3 = "Correct, Laurel's friends were not randomly assigned to each treatment. Instead, she assigned them based on their own preferences, which could lead to confounded results. For example, what if the Universe S10 testers hated Universe phones, so tried harder to damage them?"
msg4 = "Sample size is not related to experimental design. We'll talk later about how sample size matters in making inference about experimental results."
test_mc(correct = 3, feedback_msgs = c(msg1,msg2,msg3,msg4))
```

--- type:MultipleChoiceExercise lang:r xp:50 skills:1 key:1c689745b7
## Multiple Choice: Sampling
If Laurel was interested in what proportion of the U.S. population could bend a WePhone 10S at least 1 degree with their hands, which group would serve as a better sample?


*** =instructions
- Perfect strangers in Laurel's neighborhood
- Perfect strangers across the world
- Perfect strangers in Laurel's town
- Perfect strangers across the U.S.

*** =sct
```{r}
msg1 = "One neighborhood is probably not as diverse as the whole country, so try another option."
msg2 = "This answer may be tempting, but Laurel wants a sample that is reflective of the U.S. population."
msg3 = "This is close to the best case, but is unlikely to be as diverse as the whole U.S. Try again."
msg4 = "Correct. A single neighborhood is unlikely to reflect the diversity of the whole country, and involving people across the world brings in a lot of complicating confounders. People in Laurel's town is closer to ideal than those answer options, but really to get a true representation of the US population, you need to survey strangers across the US"
test_mc(correct = 4, feedback_msgs = c(msg1,msg2,msg3,msg4))
```


--- type:NormalExercise lang:r aspect_ratio:62.5 xp:50 skills:1
## Practice identifying imbalance I
Laurel has saved all of the wePhones that she has broken and she now wants to auction them off on eBay. However, Laurel is uncertain what base-auction price to set for her broken WePhones. 

To determine what price to set at the beginning of her WePhone auctions, Laurel conducts a new experiment: Laurel grabs a random sample of broken WePhones from her closet, and randomly assigns them to a treatment group (WePhones to be base priced at $10), and a control group (WePhones to be base priced at $5). Laurel's end goal is to determine whether her sample is unbalanced. Using the dataframe `WePhones`:

*** =instructions
- 1) Take a look at the structure of dataframe `WePhones` with the "str" function.
- 2) Subset the WePhones dataframe into two new dataframes, one for `WePhones` in the treatment, titled `Treatment` and one for WePhones in the control titled `Control`.
- 3) Calculate the percent of phones in the treatment group and control group that are "Black" with the mean function.

*** =pre_exercise_code
```{r}
# Initialize dataframe
    set.seed(1)
    n <- 100
    WePhones <- as.data.frame(matrix(0, ncol=6,nrow=n))
    colnames(WePhones) <- c("Treatment","Color","Hard_Drive_Capacity","Model","Condition","Network")
# Simulate baseline data
    WePhones$Treatment<-rbinom(n,1,.5)
    WePhones$Color<-rbinom(n,1,runif(1,.2,.8))
    WePhones$Hard_Drive_Capacity<-rbinom(n,1,runif(1,.2,.8))
    WePhones$Model<-rbinom(n,1,runif(1,.2,.8))
    WePhones$Network<-rbinom(n,1,runif(1,.2,.8))
    WePhones$Condition<-rbinom(n,1,runif(1,.2,.8))
# Simulate unbalanced data
  # probability baseline
      p<-runif(1,.1,.7)
  # Randomly assign unbalanced variable
      WePhones[,sample(2:6,1)]<-ifelse(WePhones$Treatment==1,rbinom(n,1,p),rbinom(n,1,p+.2))
# Transform into character variables
    WePhones$Color<-as.factor(ifelse(WePhones$Color==1,"Platinum","Black"))
    WePhones$Hard_Drive_Capacity<-as.factor(ifelse(WePhones$Hard_Drive_Capacity==1,"64gb","32gb"))
    WePhones$Model<-as.factor(ifelse(WePhones$Model==1,"8","7"))
    WePhones$Network<-as.factor(ifelse(WePhones$Network==1,"Mendizon","Trot"))
    WePhones$Condition<-as.factor(ifelse(WePhones$Condition==1,"Broken","In_Pieces"))
```
*** =sample_code
```{r}
# 1) Let's practice identifying whether the treatment and control are different for Color. First, take a look at the structure of the dataframe with the "str" function. Insert the word `WePhones` inbetween the structure function
    str() 

# 2) Notice that each variable is a factor with two levels. For ease of computation, let's subset the WePhones dataframe into two new dataframes, one for WePhones in the treatment, and one for WePhones in the control. We provide syntax for creating how to make a dataframe of just the treatment group. Use this example to create a dataframe for just the control titled `Control` 
    Treatment<-WePhones[WePhones$Treatment==1,]

# 3) Now, let's calculate the percent of phones in the treatment group and control group that are "Black". We provide an example of how to do this for the Treatment group.
    mean(Treatment$Color=="Black")

```
*** =solution
```{r}
    str(WePhones) 
    Control<-WePhones[WePhones$Treatment==0,]
    mean(Control$Color=="Black")
```
*** =sct
```{r}
test_function("str", incorrect_msg = "Did you use the `str` function?")
test_function("mean", incorrect_msg = "Did you use the `mean` function?")
test_error()

success_msg("Good work! Exploring your new data before diving into deeper questions is a critical step for conducting statistical analyses. The number of Black and White WePhones in the previous question looked different, but are they? Let's do the next question to find out!")
```


--- type:NormalExercise lang:r aspect_ratio:62.5 xp:50 skills:1
## Practice identifying imbalance
Now that we are familiar with Laurel's WePhones dataset, let's determine whether Laurel's sample is unbalanced. Using the data frame `WePhones`:

*** =instructions
- 1) Use the `t.test` function to find out if the `Treatment` and `Control` groups have the same number of `Black` WePhones.
- 2) Identify which variable in the Use the `Treatment` and `Control` groups is imbalanced.

*** =pre_exercise_code
```{r}
# Initialize dataframe
    set.seed(1)
    n <- 100
    WePhones <- as.data.frame(matrix(0, ncol=6,nrow=n))
    colnames(WePhones) <- c("Treatment","Color","Hard_Drive_Capacity","Model","Condition","Network")
# Simulate baseline data
    WePhones$Treatment<-rbinom(n,1,.5)
    WePhones$Color<-rbinom(n,1,runif(1,.2,.8))
    WePhones$Hard_Drive_Capacity<-rbinom(n,1,runif(1,.2,.8))
    WePhones$Model<-rbinom(n,1,runif(1,.2,.8))
    WePhones$Network<-rbinom(n,1,runif(1,.2,.8))
    WePhones$Condition<-rbinom(n,1,runif(1,.2,.8))
# Simulate unbalanced data
  # probability baseline
      p<-runif(1,.1,.7)
  # Randomly assign unbalanced variable
      WePhones[,sample(2:6,1)]<-ifelse(WePhones$Treatment==1,rbinom(n,1,p),rbinom(n,1,p+.2))
# Transform into character variables
    WePhones$Color<-as.factor(ifelse(WePhones$Color==1,"Platinum","Black"))
    WePhones$Hard_Drive_Capacity<-as.factor(ifelse(WePhones$Hard_Drive_Capacity==1,"64gb","32gb"))
    WePhones$Model<-as.factor(ifelse(WePhones$Model==1,"8","7"))
    WePhones$Network<-as.factor(ifelse(WePhones$Network==1,"Mendizon","Trot"))
    WePhones$Condition<-as.factor(ifelse(WePhones$Condition==1,"Broken","In_Pieces"))
    # Let's practice identifying whether the treatment and control are different for Color. First, take a look at the structure of the dataframe with the "str" function
    str(WePhones) 

# Notice that each variable is a factor with two levels. For ease of computation, let's subset the WePhones dataframe into two new dataframes, one for WePhones in the treatment, and one for WePhones in the control
    Treatment<-WePhones[WePhones$Treatment==1,]
    Control<-WePhones[WePhones$Treatment==0,]
    
# Now, let's calculate the percent of phones in the treatment group and control group that are "Black"
    mean(Treatment$Color=="Black")
    mean(Control$Color=="Black")
```
*** =sample_code
```{r}
# 1) Use the `t.test` function to find out if the `Treatment` and `Control` groups have the same number of `Black` WePhones. Replace the `x` and `y` in the following example with, `Treatment$Color=="Black"` and `Control$Color=="Black"` respectively.
    t.test(Treatment$Color=="Black",Control$Color=="Black")

# Note: The p-value derived from a t-test function computes the probability that the difference in the means between two sets of numbers (vectors) is equal to 0. These results show that there is a 0.46 (or 46%) chance that the difference in the means between the number of black phones in the treatment and control groups is equal to 0. In other words, it is fairly likely that the number of black phones in the treatment and control groups are the same. Typically, we only refute the null hypothesis (in this case, that the phone specifications are imbalanced) when there is less than a 5% chance of it being true). 

# 2) Identify which variable in the Use the `Treatment` and `Control` groups is imbalanced. Specifically, use the `t.test` function to search through each variable in the `Treatment and `Control` groups until you find one that is imbalanced (i.e. where the probability that the mean difference between the specification in the treatment and control groups is less than 5%). Write the name of the unbalanced trait between control and treatment groups into Solution1.
    t.test(  )
    Solution1<-" " 


```
*** =solution
```{r}
  a<-vector()
for(i in 2:6){
  a[i-1]<-t.test(WePhones[,i][WePhones$Treatment==0]==unique(WePhones[,i])[1],WePhones[,i][WePhones$Treatment==1]==unique(WePhones[,i])[1])$p.value
}    
  Solution1<-names(WePhones)[match(a[a<.05],a)+1]
```
*** =sct
```{r}
test_object("Solution1")

success_msg("Good work! The methods for computing balance in this exercise are very common to experimental research. While we walked you through the steps this time, some of the later problems will be less guided. You may want to return to this exercise and look at how we computed imbalance if you ever get stuck in later ones.")
```
