--- 
title       : "Experiments 1: Introduction"
description : "These videos will help you think about experiments"
  
--- type:VideoExercise lang:r aspect_ratio:62.5 xp:50 skills:1 key:4070242f87
## Introduction to the Chapter on Experiments
*** =video_link
//player.vimeo.com/video/205124502
 
--- type:VideoExercise lang:r aspect_ratio:62.5 xp:50 skills:1 key:764f98ae81
## A Quick Response
*** =video_link
//player.vimeo.com/video/205124567


--- type:MultipleChoiceExercise lang:r xp:50 skills:1 key:f571c52e35
## Multiple Choice: Rainfall in Sonoma County
Rainfall in Sonoma County, California is highly correlated with eBay's total gross merchandise volume. Which of the following answers could support a **causal explanation** for this relationship? 

*** =instructions
- Across the world, people prefer to buy more eBay products on rainy days.
- Climate change has caused Sonoma County to have increased rainfall and has encouraged the global population to do more shopping from home.
- eBay sales and rainfall in Sonoma County have both been increasing at similar rates in recent years
- Rainfall motivates residents of Sonoma County to buy and sell goods on eBay at extraordinarily high rates

*** =sct
```{r}
msg1 = "This answer is close to correct, but not quite. Remember, eBay sells its goods to people across the world---why would the correlation be so strong with rainfall just in Sonoma County? Try again"
msg2 = "This is a possible explanation for the correlation between eBay sales and rainfall in Sonoma County, but it is not a causal relationship. Instead it would suggest that this correlation is spurious. Try again"
msg3 = "This is what it means for there to be a correlation between rainfall in Sonoma County and eBay sales, but it is not a causal explanation for this correlation. Try again"
msg4 = "Good Job. As ridiculous as this sounds, it is what a causal relationship between rainfall in Sonoma County and eBay sales might suggest. Of course, residents of Sonoma County would have to be providing a huge proportion of all eBay transactions for the correlation to hold."
test_mc(correct = 4, feedback_msgs = c(msg1,msg2,msg3,msg4))
```



--- type:MultipleChoiceExercise lang:r xp:50 skills:1 key:caffaa946e
## Multiple Choice: Experimental design
Annual Apple iPhone sales are highly correlated with annual eBay sales. If this relationship is spurious, what might be causing this high correlation?

*** =instructions
- Most transactions on eBay are for Apple iPhones.
- Time causes technology sales and technology usage to increase
- Most people only shop on eBay using the iPhone's eBay app.
- Buying an iPhone makes people more comfortable with using technology for market transactions.

*** =sct
```{r}
msg1 = "This would suggest a causal relationship rather than a spurious correlation. Try again"
msg2 = "Correct! This would cause both iPhone and eBay sales to increase, but has nothing to say about a link between the two."
msg3 = "This would suggest a causal relationship rather than a spurious correlation. Try again"
msg4 = "This would suggest a causal relationship rather than a spurious correlation. Try again"
test_mc(correct = 2, feedback_msgs = c(msg1,msg2,msg3,msg4))
```
