--- 
title       : "Experiments 6: Common Issues in Experiments"
description : "These videos and exercises discuss common issues in carrying out experiments."

--- type:VideoExercise lang:r aspect_ratio:62.5 xp:50 skills:1 key:ef7f2e2846
## Important Issues in Experiment Design These Modules Ignore
*** =video_link
//player.vimeo.com/video/198212060


--- type:VideoExercise lang:r aspect_ratio:62.5 xp:50 skills:1 key:5f5a6d0023
## Other Issues in Experiments (Including Money & Ethics)
*** =video_link
//player.vimeo.com/video/209272116


--- type:MultipleChoiceExercise lang:r xp:50 skills:1 key:e302f4bddd
## Drawing Conclusions from Experiments
The transportation network company, Unter Technologies, is interested in increasing its revenue. Its CEO hypothesizes that the firm would generate substantially greater profits if it lowered its per-ride costs and profit margins. 

Unter conducts an experiment on how its sales are sensitive to price reductions by offering a 25% discount for all of its services for one day. To their pleasant surprise, Unter sees a huge spike in sales and profits during the promotional day. 

Seeing the results of this experiment, Unter's CEO decides that Unter should lower its prices permanently. Why might this conclusion be a bit hasty?

*** =instructions
- Unter did not survey its customers to determine if they were enticed by Unter's lower prices.
- Unter does not have enough data to test the statistical significance of its findings.
- The treatment in Unter's experiment might not have the same effect as the treatment in their policy proposal.
- Unter's experiment didn't address the strategic responses of its competitors.
*** =sct
```{r}
msg1 = "Doing additional research to learn The Why behind a behavioral change can be helpful, but there's an experimental design issue to address first. Try again"
msg2 = "This question doesn't say how much data Unter used, but there's another experimental design issue to consider, so try again"
msg3 = "Correct! Unter's experiment specifically tests the effect on profits of a temporary price reduction in its services, rather than the effect of a permanent price reduction. Profits increased during the promotional period, but there is no evidence that profits would remain higher over a longer period of time."
msg4 = "While Unter's profits are impacted by its competitors' behavior, there is another experimental design issue that is more appropriate. Try again"
test_mc(correct = 3, feedback_msgs = c(msg1,msg2,msg3,msg4))
```

--- type:MultipleChoiceExercise lang:r xp:50 skills:1 key:fa050e19b1
## Generalizing Experimental Results
Unter's CEO is still convinced that lowering the prices of their services might increase the company's profits. He is hesitant to lower Unter's prices for a longer period, since the company would risk losing a great deal of income if the price-drop did not substantially increase revenue. 

Therefore, to understand how a long-term price drop would affect Unter's profits, the CEO of Unter decides to drop the prices of their services in one city. Over the course of a year, Unter's profits in that city rose substantially. Unter's CEO is now convinced that he should drop Unter's prices nation-wide. Why might this conclusion still be too hasty?

*** =instructions
- Unter's experiment did not pay attention to how its changes affected its competitors.
- The context in Unter's experiment might not generalize to other contexts affected by Unter's policy proposal.
- Unter's experiment did not specify why people might be enticed by lower prices.
- Unter's experiment did not control for noncompliance.
*** =sct
```{r}
msg1 = "The company probably does exist in a complex market, but there's something else to worry about without that consideration. Try again."
msg2 = "Correct! Unter's experimental results may have been specific to the city that Unter conducted the experiment in. People from other cities may not be as sensitive to price-reductions as the city in Unter's experiment."
msg3 = "Doing some research to learn The Why behind a behavior change can be helpful, but there's an experiment design issue to address first. Try again."
msg4 = "The question assumes the CEO can successfully set prices in his own company, so if that's true, what other answer might still be a design issue we need to worry about?"
test_mc(correct = 2, feedback_msgs = c(msg1,msg2,msg3,msg4))
```

--- type:MultipleChoiceExercise lang:r xp:50 skills:1 key:7c1f83a32b
## Ethics in experiments
Which of the following causal questions is it possible to answer with a randomized experiment (without major ethical concerns)?

*** =instructions
- The effect of losing one's job on mental health.
- The effect of religion on financial well-being.
- The effect of social isolation on academic performance.
- The effect of financial incentives on athletic performance.
*** =sct
```{r}
msg1 = "This would definitely cause major ethical concerns because randomly assigning people to lose their job is not allowed.Try again."
msg2 = "This is a very personal topic to many people, and likely to involve big ethical concerns, so try again."
msg3 = "Randomly assigning people to social isolation is incredibly unethical, so look for a better answer!"
msg4 = "Correct! Each of the other experiments is not only difficult to create, but could cause substantial harm to an experiment's participants."
test_mc(correct = 4, feedback_msgs = c(msg1,msg2,msg3,msg4))
```

--- type:MultipleChoiceExercise lang:r xp:50 skills:1 key:9ca8c3a544
## Common Issues in Interpretation I: Box-and-Whisker Plots
Not only are there common issues in experiments, but there are many common ways that people misinterpret statistical results. Box-and-whisker plots (also known as box plots) are a great tool for summarizing data, but can be a source of confusion for new analysts. To refamiliarize yourself with boxplots, You can read our summary of them in the Chapter 5 exercise, "Practice working with eGulf," or on various websites, such as Wikipedia (https://en.wikipedia.org/wiki/Box_plot).

Following an experiment regarding how computer keyboard layouts affect people's typing speed, the principal investigator, Kurt Devorak, compared the hand sizes among respondents in the treatment group and control group to make sure his sample was balanced. He plotted the results in two box-and-whiskers plots (illustrated in the R workspace). Which of the following conclusions **cannot** be determined from these boxplots (or almost any boxplot)?

*** =instructions
- The maximum hand size in the treatment group is identical to the maximum hand size in the control group.
- The third quartile hand size in the treatment group is identical to the third quartile hand size in the control group.
- The median hand size in the treatment group is identical to the median hand size in the control group.
- The mean hand size in the treatment group is identical to the mean hand size in the control group.

*** =pre_exercise_code
```{r}
set.seed(1)
rtnorm <- function(n, mean, sd, min = -Inf, max = Inf){
  qnorm(runif(n, pnorm(min, mean, sd), pnorm(max, mean, sd)), mean, sd)
}
options(scipen=1)
n=200
Control<-round(rtnorm(n,7.44,sd=.5,min = 6,max=9),2)
Treatment<-Control
#increase values between median and third quartile
  Treatment<-ifelse(Treatment<quantile(Treatment,.74)& Treatment>quantile(Treatment,.51),quantile(Treatment,.74),Treatment)
#increase values between median and first quartile
  Treatment<-ifelse(Treatment<quantile(Treatment,.49)& Treatment>quantile(Treatment,.26),quantile(Treatment,.49),Treatment)
  boxplot(Treatment,Control,names=c("Treatment","Control"),main="Hand Sizes (inches) across Experimental Groups")
```

*** =sample_code
```{r}
  boxplot(Treatment,Control,names=c("Treatment","Control"),main="Hand Sizes (inches) across Experimental Groups")
```
*** =sct
```{r}
msg1 = "The maximum value in both plots is indicated by the dot above the whiskers. These are identical. Try again."
msg2 = "The third quartile in both plots is indicated by the top edge of the boxes. These are identical. Try again."
msg3 = "The median value in both plots is indicated by the thick line within the box. These are identical. Try again."
msg4 = "Correct! Box-and-whiskers plots do not show the mean values in a sample. Although these two plots have identical quartile distributions, medians, and ranges, the mean values are actually different (7.8 in the treatment and 7.5 in the control)."
test_mc(correct = 4, feedback_msgs = c(msg1,msg2,msg3,msg4))
```

---
## Let's Code: Common Issues II - Non-Normal Distributions
```yaml
type: VideoExercise
lang: r
xp: 50
skills: 1
key: 83164453d5
video_link: "//player.vimeo.com/video/279737678"
```

--- type:NormalExercise lang:r aspect_ratio:62.5 xp:50 skills:1 key:d411db8160
## Common Issues in Interpretation II: Non-Normal Distributions
Many statistical models assume that the data is normally distributed like in a bell curve, where a) most observations lie around the mean, b) the frequency of data that falls far from the mean is inversely related with its distance from the mean, and c) the distribution of observations above and below the mean are relatively identical. Often, results from experiments are non-normal, which can lead to spurious findings. Therefore, data transformations often need to be made to account for such problems.

On the TV show, Kitchen Horrors, Bordon Hannity tries to revive failing restaurants by offering advice to their management and chefs. To measure the effectiveness of Hannity's efforts, a fan measures the average reviews of one of Hannity's revived restaurants before and after he offers them advice. With these results:

*** =instructions
- 1) Find the means of the reviews before and after Bordon Hannity's advice.
- 2) Run kernal density plots of the reviews with the function plot and lines.
- 3) Determine whether the reviews are normally distributed

*** =pre_exercise_code
```{r}

library (dplyr)
library (data.table)
library (purr)

Reviews_Before<-sample(c(rep(1,23),rep(2,34),rep(3,28),rep(4,6),rep(5,34)))

Reviews_After<-sample(c(rep(1,37),rep(2,4),rep(3,26),rep(4,44),rep(5,14)))
```
*** =sample_code
```{r}
# 1) It is almost always best practice to explore the data before running any treatment tests. However, let's try to be bad statisticians and dive into the results before we explore the data. The two review variables are measured on Likert scales, ranging from 1 to 5, where 1 signals that the reviewers strongly disliked the restaurant and 5 signals that reviewers strongly liked the restaurant. Let's examine the means of these reviews. Use the mean command to find the averages of the variables Reviews_Before and Reviews_After.

#---- Question 1a-------------------------------------#
    Solution1a<-
#----------------------------------------------------#

# 2) The mean reviews before and after treatment are identical! But does this mean that Hannity's show had no effect on this restaurant's ratings? Not necessarily. To illustrate, let's plot the kernel densities of the reviews in the treatment and control groups to see how they are distributed. Use the plot(density()) functions on the variables Reviews_Before and Reviews_After.

#---- Question 1b-------------------------------------#
    Solution1b<-
#----------------------------------------------------#

# 3) Use the Previous Plot and Next Plot buttons to cycle between the two graps. Do these these look approximately normally distributed? Answer with "Yes" or "No."
#---- Question 1c-------------------------------------#
    Solution1c<-""   
#----------------------------------------------------#
 
```
*** =solution
```{r}
  Solution1a<-mean(Reviews_Before)
              mean(Reviews_After)
  Solution1b<-plot(density(Reviews_Before))
              plot(density(Reviews_After))
  Solution1c<-"No"
```
*** =sct
```{r}
test_object("Solution1a")
test_object("Solution1b")
test_object("Solution1c")
test_function("mean", incorrect_msg = "Did you use the `mean` function?")
test_function("plot(density())", incorrect_msg = "Did you use the `plot(density())` function?")
test_error()
success_msg("Good work! This is an easy call in this particular case, but if it's close then you'll want to try some statistical tests too. And it might provoke some healthy discussion with your colleagues!").
```


--- type:NormalExercise lang:r aspect_ratio:62.5 xp:50 skills:1 key:5b26b74b8c
## Common Issues in Interpretation II: Non-Normal Distributions (2)
There are many ways to deal with data that are multimodal in order to get them to be normally distributed. A common strategy for dealing with data that are ordered on a Likert scale is to collapse the response options. This is feasible if we assume that it is not important to differentiate the degree of positive or negative feelings for a restaurant. In this case, we would want to collapse response options 1 and 2 into a single response option to represent negative opinions, and 4 and 5 into a single response option to represent positive opinions. 

We will create for you a new variable that collapses the Reviews_Before 5 point response scale into 3 points: negative, ambivalent, and positive, and you will do 

*** =instructions
- 1) Create a new variable that collapses Reviews_After variable into 3 points: negative, ambivalent, and positive.
- 2) Run a t-test that compares the treatment and control with the new ratings scale, and look at the p value.

*** =pre_exercise_code
```{r}

library (dplyr)
library (data.table)
library (purr)

Reviews_Before<-sample(c(rep(1,23),rep(2,34),rep(3,28),rep(4,6),rep(5,34)))

Reviews_After<-sample(c(rep(1,37),rep(2,4),rep(3,26),rep(4,44),rep(5,14)))
```
*** =sample_code
```{r}
# Let's create a new variable with three parameters, where 1= Negative (1 or 2), 2= Ambivalent (3), and 3= Positive (4 or 5). To do this for the Reviews_Before variable, we can use the ifelse function. The first line of the below syntax says, if Reviews_Before is equal to 1 or 2, set Reviews_Before to 1, otherwise just return Reviews_Before. The following lines can be interpreted similarly.
    Reviews_Before<-ifelse(Reviews_Before==1 | Reviews_Before==2,1,Reviews_Before) 
    Reviews_Before<-ifelse(Reviews_Before==3,2,Reviews_Before) 
    Reviews_Before<-ifelse(Reviews_Before==4 | Reviews_Before==5,3,Reviews_Before)
  
# 1) Use the above syntax as a guide to rescale the Reviews_After variable, so that Reviews_After is coded as 1 if it previously was coded as 1 or 2, 2 if it was previously coded as 3, and 3 if it was previously coded as 4 or 5.
#---- Question 2a-------------------------------------#
    Reviews_After<-
#----------------------------------------------------#
    
# 2): Now that you have rescaled the Reviews_After variable, run a t-test between the Reviews_After and Reviews_Before variables, and return the p-value (use t.test()$p.value)
#---- Question 2b-------------------------------------#
    Solution2b<-
#----------------------------------------------------#

```
*** =solution
```{r}
  Solution2a<-
    Reviews_Before<-ifelse(Reviews_Before==1 | Reviews_Before==2,1,Reviews_Before) 
    Reviews_Before<-ifelse(Reviews_Before==3,2,Reviews_Before) 
    Reviews_Before<-ifelse(Reviews_Before==4 | Reviews_Before==5,3,Reviews_Before)
    Reviews_After<-ifelse(Reviews_After==1 | Reviews_After==2,1,Reviews_After) 
    Reviews_After<-ifelse(Reviews_After==3,2,Reviews_After) 
    Reviews_After<-ifelse(Reviews_After==4 | Reviews_After==5,3,Reviews_After)

  Solution2b<-t.test(Reviews_After,Reviews_Before)$p.value
```
*** =sct
```{r}
test_object("Solution2a")
test_object("Solution2b")
test_function("ifelse", incorrect_msg = "Did you use the `ifelse` function?")
test_function("t.test", incorrect_msg = "Did you use the `t.test` function?")
test_error()
success_msg("Good work! If we assume that the difference in respondent's degrees of positive or negative feelings towards the restaurant are non-meaningful, we can conclude that Hannity's restaurant revival is effective (i.e. it increases the number of postive versus negative reviews). If the degree of positive or negative feelings towards restaurants are meaningful, then we do not have enough information to reject the null hypothesis (i.e. evidence does not support the claim that Hannity's contributions leads restaurants to get better reviews").
```

--- key:2cff43ba90
## Let's Code: Common Issues III - Outliers
```yaml
type: VideoExercise
lang: r
xp: 50
skills: 1
key: 
video_link: //player.vimeo.com/video/
```

--- type:NormalExercise lang:r aspect_ratio:62.5 xp:50 skills:1 key:12e1bc32a2
## Common Issues in Interpretation III: Outliers
Outliers are another common source of causal misinpretation that often confound results. An outlier is an observation that has a very distant parameter or outcome from other observations. In this question, we will learn how to identify and handle outliers. 

In a previous exercise, we examined an experiment conducted by the popular online auctioneer, eGulf, where a random sample of dedicated sellers of used WePhones was drawn. This sample only included eGulf sellers who typically post 10 pictures of their used WePhones. As an experimental treatment, eGulf temporarily allowed some of these sellers to post up to 15 pictures for each auction of their WePhones, and measured whether posting more than 10 pictures affected final WePhone sales prices. In this exercise, eGulf conducted their experiment again. Their sample size seems sufficiently large (N=350), and the mean sales prices of users in the treatment group also seem substantially higher than in the control group. But are they? Follow the code in the workspace to identify and correct for outliers.

*** =instructions
- Create boxplots of the final sales prices for the treatment and control groups.
- Identify and remove the outliers from the sample.
- Examine the mean difference between the Treatment and Control group.

*** =pre_exercise_code
```{r}

library (dplyr)
library (data.table)
library (purr)

rtnorm <- function(n, mean, sd, min = -Inf, max = Inf){
  qnorm(runif(n, pnorm(min, mean, sd), pnorm(max, mean, sd)), mean, sd)
}
set.seed(1)
n <- 350
eGulf <- as.data.frame(matrix(0, ncol=2,nrow=n))
colnames(eGulf) <- c("Seller_Feedback_Count","Seller_Feedback_Score")
# Simulate baseline data
  # Make seller feedback count
    eGulf$Seller_Feedback_Count<-as.integer(abs(round(rnorm(n,200,100))))
  # Make feedback score loosely correlated with feedback count 
    eGulf$Seller_Feedback_Score<-round(.9+.1*eGulf$Seller_Feedback_Count/max(eGulf$Seller_Feedback_Count)-rbeta(n,1,10),2)
  # Make final sales price also correlated with seller_feedback_score
    eGulf$Final_Sales_Price<-round(rlnorm(n, meanlog=log(400), sdlog = log(1.2))*eGulf$Seller_Feedback_Score)
  # Add parameters for new problem
    eGulf$Treatment<-rbinom(n,1,.5)
    eGulf$temp<-rnorm(n)
  #Create 3 outliers
    eGulf$Treatment[1:3]<-1
    eGulf$Final_Sales_Price[1:3]<-round(rtnorm(3,mean=2000,sd=200,min=1500,max=2500))
    eGulf<-eGulf[order(eGulf$temp),]
    
```
*** =sample_code
```{r}
# Although it's generally recommended to explore your data before running any statistical tests on it, for the sake of learning, let's naively estimate the average treatment effect on final sales prices with a t-test.
    t.test(eGulf$Final_Sales_Price[eGulf$Treatment==1],eGulf$Final_Sales_Price[eGulf$Treatment==0])

# The ATE for the above t-test is almost $36, but the results are not statistically significant. This seems suspicious; usually a large means difference found in a sample of this size would be statically significant. Let's examine the distribution of the treatment and control groups with a box plot. Fill in the box plot with the final sales prices for the treatment and control groups. For reference, the syntax for making this boxplot is identical to that of a t-test.
#---- Question 1-------------------------------------#
    Solution1<- boxplot()
#----------------------------------------------------#
    
# Notice that there are a few dots above each box plot. These dots indicate outliers. The cutoff for outliers in the boxplot is any value that is 1.5 times the interquartile range above the upper quartile or below the lower quartile. Is it possible that these outliers are driving our treatment effect? Let's correct for these extreme outliers. Subset the data so that it excludes any WePhone sales that were above $1500. 
#---- Question 2-------------------------------------#
    eGulf<-eGulf[]
#----------------------------------------------------#

# Question 3: Now that you have subsetted the data, calculate the mean difference in final sales prices between the treatment and control groups.
#---- Question 3-------------------------------------#
    Solution3<-
#----------------------------------------------------#

```
*** =solution
```{r}
Solution1<- boxplot(eGulf$Final_Sales_Price[eGulf$Treatment==1],eGulf$Final_Sales_Price[eGulf$Treatment==0])
eGulf<-eGulf[eGulf$Final_Sales_Price<=1500,]
Solution3<-mean(eGulf$Final_Sales_Price[eGulf$Treatment==1])-mean(eGulf$Final_Sales_Price[eGulf$Treatment==0])
```
*** =sct
```{r}
test_object("Solution1")
test_object("eGulf")
test_object("Solution3")
success_msg("Good work! When you removed the outliers from the data, the mean difference was much smaller between the treatment and control group. This is a problem that comes up often in data analysis.")
```
